### 2025-10-01 manuscript/enar/enar_cls.tex

\documentclass[referee,12pt,usenatbib]{biom}

\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsmath}
%\RequirePackage{,,}
\bibliographystyle{biom} % Biometrics bib style if installed
%\usepackage[authoryear,round]{natbib}
\title{Hybrid Partial Least Squares Regression with Multiple Functional and Scalar Predictors}

\begin{document}
	\begin{abstract}
	Motivated by renal imaging data with accompanying side information, we propose a hybrid partial least squares (PLS) regression method to capture cross-modality correlations in settings with both functional and scalar predictors. Our approach extends the nonlinear iterative PLS (NIPALS) algorithm to a hybrid Hilbert space, embedding functional and scalar predictors into a unified vector representation and iteratively maximizing their empirical cross-covariance with the response. The resulting low-dimensional representations effectively capture within- and between-modality variation and predictor-response correlation. Out method  preserves the desirable theoretical properties of PLS. Moreover, the procedure is computationally efficient, requiring only the solution of linear systems at each iteration.
\end{abstract}

\begin{keywords}
Dimension reduction;
Functional data analysis;
Multiple data modalities;
Multivariate data analysis;
Multivariate functional data;
Partial least square
\end{keywords}

\maketitle
\newpage
\section{Introduction} \label{sec: Introduction}
Biomedical studies often collect multi-modal data.
Our running example is the Emory University renal study \citep{changBayesianLatentClass2020, jangPrincipalComponentAnalysis2021}, which records two renogram curves (functional data) along with  side information (Euclidean vector) for each kidney.
Such distinct yet related physiological signals can be represented as functional- and scalar-valued i.i.d. covariates in a linear regression model:
\begin{equation}	\label{eq:hybrid_regression_model}
	Y_i
	= 
	\boldsymbol{\beta}^\top \mathbf{Z}_i
	+ 
	\sum 
	\limits_{k=1}^K 
	\int  \beta_k(t) X_{ik}(t) dt + \epsilon_i,
	\quad
	i = 1, \ldots, n,
	%= \langle \boldsymbol{\alpha}, \mathbf{Z} \rangle +  \langle \beta, X \rangle_{\mathcal{F}}  + \epsilon
\end{equation}
where $Y_i$ is a scalar response, 
$X_{i1}(t), \ldots, X_{iK}(t) \in \mathbb{L}^2[0,1]$ are functional predictors, 
$\mathbf{Z}_i = (Z_{i1}, \ldots, Z_{ip})^\top$ is a Euclidean vector covariate,
and $\epsilon_i$ denotes observational noise.
For notational convenience, we assume throughout that the responses and predictors are centered, so the intercept term can be omitted.
The key   challenge here is the strong correlations between functional and scalar predictors,
in addition to the ill-posedness of the infinite-dimensional slope functions $\beta_k(t)$ and the high dimensionality of the predictors ($K + p$). This paper addresses all three issues in a unified manner by introducing a hybrid partial least squares (PLS) regression framework defined on a novel hybrid Hilbert space.

	\paragraph{Previous works and limitations}
Basis expansion with regularization of roughness addresses the ill-posedness of functional data by exploiting their smoothness \citep{cardotSplineEstimatorsFunctional2003,aguileraPenalizedVersionsFunctional2016, 
	caiPredictionFunctionalLinear2006, zhaoWaveletbasedLASSOFunctional2012}.
High dimensionality can be addressed by using derived inputs, with PCA regression as a straightforward example. We can apply PCA  separately to functional (e.g., \citealt{happ_multivariate_2018}) and scalar predictors, followed by multivariate regression on the combined scores. While well studied for functional predictors alone \citep{hallMethodologyConvergenceRates2007, reissFunctionalPrincipalComponent2007, febrero-bandeFunctionalPrincipalComponent2017}, PCA regression may fail to capture the core regression relationship since the derived inputs are not guided by the response.
Partial least squares (PLS) regression is a powerful alternative that constructs orthogonal latent components from predictors to maximize covariance with the response.
For functional predictors, \citet{predaPLSRegressionStochastic2005} introduced PLS regression for a single predictor, with subsequent extensions incorporating
basis approximations and roughness penalties
\citep{reissFunctionalPrincipalComponent2007, aguileraUsingBasisExpansions2010, aguileraPenalizedVersionsFunctional2016}, and multiple functional predictors \citep{beyaztasRobustFunctionalPartial2022}.
Theoretical and computational advances were developed by \citet{delaigleMethodologyTheoryPartial2012} and \citet{saricamPartialLeastsquaresEstimation2022}, respectively; see \citet{febrero-bandeFunctionalPrincipalComponent2017} for a review.
These approaches ignore correlations between functional and scalar components, leading to multicollinearity and unreliable prediction. Multimodal correlations have mainly been explored in unsupervised settings,
such as graphical models, precision matrix estimation, and joint PCA \citep{kolar_graph_2014, gengJointNonparametricPrecision2020, jangPrincipalComponentAnalysis2021}. However,  these methods are response-agnostic and may miss outcome-relevant correlations.

\paragraph{Our contributions}
To address these gaps, we propose a hybrid PLS regression framework that unifies basis expansion, PLS, and functional-scalar correlations. Treating functional and Euclidean components as a single vector in a Hilbert space with a suitable inner product, we iteratively extract joint directions that maximize covariance with the response under a unit-norm constraint. The framework handles dense or irregular functional data, supports regularization to reduce overfitting, and comes with theoretical guarantees.

			\bibliography{bibliography.bib}

\end{document}