### 2025-10-08 create-FSHybridPLS.Rmd
hybrid predictor object now has eval_point

---
title: "Creating the ``r params$package_name`` R package"
author: "Your Name"
date: "The Date"
knit: litr::render
output: litr::litr_html_document
params:
  package_name: "FSHybridPLS" # <-- change this to your package name
  package_parent_dir: "." # <-- relative to this file's location
---

<!-- This Rmd file contains all the code needed to define an R package.  Press "Knit" in RStudio or more generally run `litr::render("name-of-this-file.Rmd")` to generate the R package.  Remember that when you want to modify anything about the R package, you should modify this document rather than the package that is outputted.
-->

## Package setup

We start by specifying the information needed in the DESCRIPTION file of the R package.
 library()
library(fda)
```{r package-setup, message=FALSE, results='hide'}
usethis::create_package(
  path = ".",
  fields = list(
    Package = params$package_name,
    Version = "0.0.0.9000",
    Title = "A Package That Says Hello",
    Description = "This package says hello.  But its actual purpose is to show how an R package can be completely coded in a single R markdown file.",
    Imports = "fda (>= 6.1.3), Matrix, funData, foreach, irlba, refund, R6",
    `Authors@R` = person(
      given = "First",
      family = "Last",
      email = "you@gmail.com",
      role = c("aut", "cre")
      )
  )
)
usethis::use_mit_license(copyright_holder = "F. Last")
```

# helper functions

### is_same_basis
```{r}
#' @export
is_same_basis <- function(input, other) {
  all(
    length(input$functional_list) == length(other$functional_list),
    all(
      mapply(function(fd1, fd2) fda::is.eqbasis(fd1$basis, fd2$basis),
             input$functional_list, other$functional_list)
    )
  )
}
```

### compute_gram_matrix
```{r}
#' Compute the Gram matrix for a basis object
#'
#' @param basis A basis object of class `basisfd` from the `fda` package
#' @return A square matrix where entry (i,j) is the inner product of basis functions i and j
#' @export
compute_gram_matrix <- function(basis) {
  stopifnot(inherits(basis, "basisfd"))
  inprod(basis, basis)  # computes ∫ b_i(t) b_j(t) dt
}

```

```{r}
#' Check if all matrices in a list are identical (with tolerance for numeric comparison)
#'
#' This function iterates through a list of matrices and checks if all subsequent
#' matrices are numerically identical to the first matrix in the list.
#' It uses `all.equal` for robust comparison of numeric matrices, which accounts
#' for small floating-point differences.
#'
#' @param gram_list A list of matrices.
#' @param tolerance A numeric tolerance for comparing floating-point numbers.
#'                  Defaults to `sqrt(.Machine$double.eps)`.
#' @return A logical value: `TRUE` if all matrices are identical, `FALSE` otherwise.
#' @export
are_all_gram_matrices_identical <- function(gram_list, tolerance = sqrt(.Machine$double.eps)) {
  # Handle edge cases: empty list or list with a single matrix
  if (length(gram_list) == 0) {
    message("The input list is empty. Returning TRUE as there are no differences.")
    return(TRUE)
  }
  if (length(gram_list) == 1) {
    message("The input list contains only one matrix. Returning TRUE.")
    return(TRUE)
  }

  # Get the first matrix as the reference
  first_matrix <- gram_list[[1]]

  # Ensure the first element is a matrix
  if (!is.matrix(first_matrix)) {
    stop("The first element of 'gram_list' is not a matrix.")
  }

  # Iterate from the second matrix onwards and compare with the first
  for (i in 2:length(gram_list)) {
    current_matrix <- gram_list[[i]]

    # Ensure current element is a matrix
    if (!is.matrix(current_matrix)) {
      stop(paste("Element", i, "of 'gram_list' is not a matrix."))
    }

    # Compare the current matrix with the first matrix
    # all.equal returns TRUE if identical, or a character string describing differences
    # We convert the result to a logical TRUE/FALSE
    if (!isTRUE(all.equal(first_matrix, current_matrix, tolerance = tolerance))) {
      # If not equal, return FALSE immediately
      message(paste("Matrices at index 1 and", i, "are not identical."))
      return(FALSE)
    }
  }

  # If the loop completes, all matrices are identical
  return(TRUE)
}
```

## helpers for functional data objects

### rep_fd

```{r}
#' Replicate a list of single-sample fd objects multiple times
#'
#' This function broadcasts a list of single-sample functional predictors
#' by replicating their coefficient columns.
#'
#' @param fd_list A list of `fd` objects, each representing a single sample.
#' @param n The number of replications desired.
#'
#' @return A list of `fd` objects, each with `n` replications.
#' @export
rep_fd <- function(fd_list, n) {
  if (!is.list(fd_list) || any(!vapply(fd_list, inherits, logical(1), "fd"))) {
    stop("Input must be a list of 'fd' objects.")
  }

  lapply(fd_list, function(fd_obj) {
    coef_mat <- fd_obj$coefs
    if (is.null(dim(coef_mat)) || ncol(coef_mat) != 1) {
      stop("Each fd object in the list must have one column of coefficients.")
    }

    new_coefs <- matrix(rep(coef_mat, n), nrow = nrow(coef_mat), ncol = n)
    fd(coef = new_coefs, basisobj = fd_obj$basis)
  })
}
```

| Case | Description                            | Input Type               | Expected Output                        | Notes                                            |
|------|----------------------------------------|--------------------------|----------------------------------------|--------------------------------------------------|
| 1    | Correct replication                    | List of single-sample fd | List of `fd` objects with `n` columns | All columns identical to original coefficients   |
| 2    | Each fd has one sample                 | List of fd               | No error                               | Verifies structure and content                  |
| 3    | One fd has multiple samples            | fd with >1 columns       | Error                                   | Should raise informative error                  |
| 4    | Non-fd object in list                  | List of non-`fd` items   | Error                                   | Validates class-checking robustness             |

# Class predictor_hybrid
- Let $\{X^{(k)}\}_{k=1, \ldots, K}$ be a collection of random functions defined on unit hypercube domains $\mathcal{T}_k := [0,1]^{d_k}$ ($d_k \in \mathbb{N})$; i.e., $X^{(k)}: \mathcal{T}_k \rightarrow \mathbb{R}$.

- Assume that each $X^{(k)}$ is in $L^2(\mathcal{T}_k)$, a Hilbert space of square integrable functions with respect to Lebesgue measure $dt_k$ on $\mathcal{T}_k$.

- Write  $X=(X^{(1)}, \ldots, X^{(K)})$ as a multivariate functional object that belongs to $\mathcal{F} = L^2(\mathcal{T}_1) \times \cdots \times L^2(\mathcal{T}_K)$---a cartesian product of individual $L^2(\mathcal{T}_k)$ spaces.

- We can also express the functional object $X$ evaluated on the multi-dimensional argument $\mathbf{t} = (t_1, \ldots, t_k)^\top \in \mathcal{T} = \mathcal{T}_1 \times \cdots \times \mathcal{T}_K$ as a $K$-dimensional vector $X(\mathbf{t})=(X^{(1)}(t_1), \ldots, X^{(K)}(t_K))^\top$.

- **Our strategy is to formulate a hybrid random object**, $\mathbf{W} = (X, \mathbf{Z})$, which combines $X$ and $p$-dimensional scalar covariate $\mathbf{Z}$ into an ordered pair belonging to $\mathcal{H} = \mathcal{F} \times \mathbb{R}^p$.

- An alternative notation for the hybrid object can be obtained by evaluating its functional part on $\mathbf{t}$ and expressing it as a $(K+p)$-dimensional vector: $\mathbf{W}[\mathbf{t}] = (X(\mathbf{t}), \mathbf{Z})^\top$, with $X(\mathbf{t}) = (X^{(1)}(t_1), \ldots, X^{(K)}(t_K))^\top \in \mathbb{R}^K$ and $\mathbf{Z} = (Z_1, \ldots, Z_p)^\top  \in \mathbb{R}^p$.

- The hybrid object forms a Hilbert space, characterized by well-defined addition, scalar multiplication, and inner product operations.

- We define an S3 object `predictor_hybrid` that represent this structure, and implement Hilbert space operations for this class, and our algorithm is built on these foundational operations.

The class definition is a simple list that contains a matrix and a collection of \code{fd} objects from the \code{fda} package, along with the necessary metadata.
- gram matrices are saved as a list.
```{r classdef:predictor_hybrid}
#' Create a predictor_hybrid object (S3 version, automatic basis size and Gram matrix)
#' Constructs a hybrid predictor object that stores both scalar and functional predictors,
#' and automatically computes Gram matrices (inner products of basis functions).
#'
#' @param Z A numeric matrix of dimension \code{n_sample × n_scalar} representing scalar predictors.
#' @param functional_list A list of functional predictors (e.g., \code{fd} objects from the \code{fda} package).
#'
#' @return An object of class \code{predictor_hybrid}, containing scalar and functional data with Gram matrices.
#' @export
predictor_hybrid <- function(Z, functional_list, eval_point) {
  stopifnot(is.matrix(Z), is.numeric(Z))
  stopifnot(is.list(functional_list))

  n_sample <- nrow(Z)
  n_scalar <- ncol(Z)
  n_functional <- length(functional_list)

  gram_list <- gram_deriv2_list <- vector("list", n_functional)
  n_basis_list <- numeric(n_functional)

  for (i in seq_len(n_functional)) {
    fd_i <- functional_list[[i]]
    stopifnot(inherits(fd_i, "fd"))

    n_fd_sample <- ncol(coef(fd_i))
    if (n_fd_sample != n_sample) {
      stop(sprintf("Functional predictor %d has %d replicates, but Z has %d rows.",
                   i, n_fd_sample, n_sample))
    }

    basis_i <- fd_i$basis
    gram_list[[i]] <- compute_gram_matrix(basis_i)
    gram_deriv2_list[[i]] <- fda::getbasispenalty(basis_i) # New

    n_basis_list[i] <- basis_i$nbasis
  }

  structure(
    list(
      Z = Z,
      functional_list = functional_list,
      eval_point = eval_point,
      gram_list = gram_list,
      gram_deriv2_list = gram_deriv2_list, # Store the new list
      n_basis_list = n_basis_list,
      n_sample = n_sample,
      n_functional = n_functional,
      n_scalar = n_scalar
    ),
    class = "predictor_hybrid"
  )
}
```

### predictor_hybrid_from_coef
An alternative constructor for a predictor_hybrid object based on a flat numeric coefficient vector. It reconstructs both the functional and scalar components for a single-sample hybrid predictor object. The function reads the basis information from the template `format` object.

- Inputs
  - **`format`**: A `predictor_hybrid` object containing:
    - `functional_list`: List of `fd` objects with basis info
    - `gram_list`: List of Gram matrices
    - `n_scalar`: Number of scalar predictors
  - **`coef`**: Numeric vector of length `K * M + d`, where:
    - `K` = number of functional predictors  
    - `M` = number of basis functions (assumed same across predictors)  
    - `d` = number of scalar predictors

- Output
  - A `predictor_hybrid` object with:
    - Functional predictors rebuilt from the first `K * M` coefficients
    - Scalar predictor `Z` from the remaining `d` coefficients
    - `n_sample` set to 1

```{r}
#' Construct a Single-Sample Predictor Hybrid Object from Coefficients
#' Reconstructs a \code{predictor_hybrid} object representing one observation, using a numeric
#' coefficient vector. This alternative constructor maps the coefficients back into their functional
#' and scalar predictor representations based on the structure of a template \code{predictor_hybrid} object.
#'
#' @param format A \code{predictor_hybrid} object that provides the structure and basis information.
#' @param coef A numeric vector containing coefficients for both functional and scalar predictors.
#'
#' @return A \code{predictor_hybrid} object with updated \code{functional_list}, \code{Z}, and \code{n_sample = 1}.
#'
#'
#' @export
predictor_hybrid_from_coef <- function(format, coef){
  M <- format$n_basis_list[1] # number of basis functions
  K <- format$n_functional # number of functional predictors
  for (ii in 1:K){
    format$functional_list[[ii]] <- fd(
      coef = as.matrix(coef[((ii-1)*M + 1):(ii*M)]),
      basisobj = format$functional_list[[ii]]$basis
    )
  }
  format$Z <- t(as.matrix(coef[(K * M + 1):length(coef)]))
  format$n_sample <- 1
  return(format)
}
```

## Basic arithmetic

### add.predictor_hybrid
Performs element-wise addition of two `predictor_hybrid` objects.  Functional predictors are combined using `plus.fd()` and `times.fd()` from the `fda` package.

- Usage

```r
add.predictor_hybrid(input, other, alpha = 1)
```

- Arguments:

  - `input`: A `predictor_hybrid` object.
  - `other`: Another `predictor_hybrid` object to be added.
  - `alpha`: A scalar multiplier applied to `other` before addition (default is 1).

- Value:
Returns a new `predictor_hybrid` object representing the result of the addition.

- Details: 
  - This function assumes both objects have the same number and structure of functional and scalar predictors.  
  - Functional parts are scaled by `alpha` using `times.fd()` and then summed using `plus.fd()`.  
  - Scalar predictors are added using standard matrix addition.

```{r}
add.predictor_hybrid <- function(xi_1, xi_2, alpha = 1) {
  # Safe access to is.eqbasis()
  is_eqbasis <- getFromNamespace("is.eqbasis", "fda")

  for (i in seq_len(xi_1$n_functional)) {
    if (!is_eqbasis(xi_1$functional_list[[i]]$basis, xi_2$functional_list[[i]]$basis)) {
      stop("Functional predictors must have the same basis.")
    }
  }

   # Type checks
  if (!inherits(xi_1, "predictor_hybrid") || !inherits(xi_2, "predictor_hybrid")) {
    stop("Both inputs must be of class 'predictor_hybrid'.")
  }

  # Structural checks
  if (xi_1$n_functional != xi_2$n_functional) {
    stop("Mismatch in number of functional predictors.")
  }
  if (xi_1$n_scalar != xi_2$n_scalar) {
    stop("Mismatch in number of scalar predictors.")
  }

  # Swap so that broadcasting always applies to xi_2
  if (xi_1$n_sample == 1 && xi_2$n_sample > 1) {
    tmp <- xi_1
    xi_1 <- xi_2
    xi_2 <- tmp
  }

  n1 <- xi_1$n_sample
  n2 <- xi_2$n_sample

  if (!(n1 == n2 || n2 == 1)) {
    stop("Sample sizes are incompatible for broadcasting.")
  }

  # Prepare components
  f1 <- xi_1$functional_list
  f2 <- xi_2$functional_list
  Z1 <- xi_1$Z
  Z2 <- xi_2$Z

  # Replicate fd and Z if needed
  if (n2 == 1) {
    f2 <- rep_fd(f2, n1)
    Z2 <- matrix(rep(c(Z2), n1), nrow = n1, byrow = TRUE)
  }

  # Combine functional predictors
  new_functional_list <- Map(
    function(f1, f2) plus.fd(f1, times.fd(alpha, f2)),
    f1,
    f2
  )

  # Compute scalar inner products
  inprod_scalar <- rowSums(Z1 * Z2) 
  # Combine scalar predictors
  new_Z <- Z1 + alpha * Z2

  # Construct new object
  predictor_hybrid(Z = new_Z, functional_list = new_functional_list)
}
```

### subtr.predictor_hybrid
Performs element-wise subtraction of two `predictor_hybrid` objects.   Internally uses `add.predictor_hybrid(input, other, alpha = -1)` to compute the result by negating the second operand.

- Usage

```r
subtr.predictor_hybrid(input, other, alpha = 1)
```

- Arguments:

  - `input`: A `predictor_hybrid` object.
  - `other`: Another `predictor_hybrid` object to be subtracted.
  - `alpha`: A scalar multiplier applied to `other` before subtraction (default is 1).

- Value:  
Returns a new `predictor_hybrid` object representing the result of the subtraction.

- Details:  
  - This function assumes both objects have the same number and structure of functional and scalar predictors.  
  - It performs subtraction by internally calling `add.predictor_hybrid()` with `-alpha`.  
  - Functional parts are scaled using `times.fd()` and subtracted via `plus.fd()` with a negated factor.  
  - Scalar predictors are subtracted using standard matrix arithmetic.

```{r}
#' Subtract two predictor_hybrid objects
#'
#' Performs element-wise subtraction of two `predictor_hybrid` objects.
#' Internally uses `add.predictor_hybrid(input, other, alpha = -1)`.
#'
#' @param input A `predictor_hybrid` object.
#' @param other Another `predictor_hybrid` object to subtract.
#' @param alpha A scalar multiplier applied to `other` before subtraction (default is 1).
#'
#' @return A new `predictor_hybrid` object representing the result of subtraction.
#' @export
subtr.predictor_hybrid <- function(input, other, alpha = 1) {
  add.predictor_hybrid(input, other, alpha = -alpha)
}

```
We skip the unit test for this function.

### scalar_mul.predictor_hybrid
Multiplies all components of a `predictor_hybrid` object by a scalar. Functional components are scaled using `times.fd()` from the `fda` package, and scalar predictors are multiplied directly using matrix operations.

- Usage

```r
scalar_mul.predictor_hybrid(input, scalar)
```

- Arguments:

  - `input`: A `predictor_hybrid` object.
  - `scalar`: A numeric value used to scale both scalar and functional components.

- Value:  
Returns a new `predictor_hybrid` object with all components scaled by `scalar`.

- Details:  
  - Functional predictors are scaled using `times.fd(scalar, fd_obj)` for each element.  
  - Scalar predictors (the matrix `Z`) are scaled elementwise using matrix multiplication.

```{r}
#' Multiply a predictor_hybrid object by a scalar
#'
#' Performs scalar multiplication on both the scalar and functional components
#' of a `predictor_hybrid` object. Functional predictors are scaled using 
#' `times.fd()` from the `fda` package.
#'
#' @param input A `predictor_hybrid` object.
#' @param scalar A numeric value to multiply all components by.
#'
#' @return A new `predictor_hybrid` object scaled by `scalar`.
#' @export
scalar_mul.predictor_hybrid <- function(input, scalar) {
  if (!inherits(input, "predictor_hybrid")) {
    stop("Input must be of class 'predictor_hybrid'.")
  }
  if (!is.numeric(scalar) || length(scalar) != 1) {
    stop("Scalar must be a single numeric value.")
  }

  # Scale functional components
  new_functional_list <- lapply(input$functional_list, function(fd_obj) {
    times.fd(scalar, fd_obj)
  })

  # Scale scalar predictors
  new_Z <- scalar * input$Z

  # Reconstruct hybrid object
  predictor_hybrid(
    Z = new_Z,
    functional_list = new_functional_list
  )
}
```

### inprod.predictor_hybrid
- The inner product of $f_1=(f_1^{(1)}, \ldots,f_1^{(K)})$ and $f_2=(f_2^{(1)}, \ldots,f_2^{(K)})$ in $\mathcal{F}$ is defined as 
$$\langle f_1, f_2\rangle_\mathcal{F} =  \sum_{k=1}^K \langle f_1^{(k)}, f_2^{(k)}\rangle_{L^2} = \sum_{k=1}^K \int_{\mathcal{T}_k} f_1^{(k)}(t_k) f_2^{(k)}(t_k) dt_k,$$ with norm 
$$\Vert f_1 \Vert_\mathcal{F} = \langle f_1,f_1 \rangle_\mathcal{F}^{1/2} = \{ \sum_{k=1}^K \int_{\mathcal{T}_k} f_1^{(k)}(t_k)^2 dt_k\}^{1/2}.$$
- We define the inner product between any two hybrid objects, $\mathbf{h}_1 = (f_1, \mathbf{v}_1)$ and $\mathbf{h}_2 = (f_2, \mathbf{v}_2)$, as
$$
    \langle \mathbf{h}_1, \mathbf{h}_2\rangle_{\mathcal{H}} = \langle f_1, f_2\rangle_\mathcal{F} +  \langle  \mathbf{v_1}, \mathbf{v_2} \rangle = \sum \limits_{k=1}^K \int_{\mathcal{T}_k} f_1^{(k)}(t_k) f_2^{(k)}(t_k) dt_k + \sum \limits_{r=1}^p v_{1r}v_{2r}, 
$$
with norm $\Vert \cdot \Vert_\mathcal{H} = \langle \cdot,\cdot \rangle_\mathcal{H}^{1/2}$.

- The method `inprod.predictor_hybrid` computes the inner product between two `predictor_hybrid` objects. It supports broadcasting when one of the inputs has a single observation.

- Usage

```r
inprod.predictor_hybrid(xi_1, xi_2)
```

- Arguments:

  - `xi_1`: A `predictor_hybrid` object.
  - `xi_2`: Another `predictor_hybrid` object. If omitted, defaults to xi_1 (computes self-inner product).

- Value:  A numeric vector of inner products (or a scalar if both inputs are single observations).

- Details:  
  - Functional components are summed using inprod() from the fda package.
  - Scalar components are handled via matrix multiplication.
  - Broadcasting is supported: if either xi_1 or xi_2 has only one sample, its values are broadcast across all rows of the other.
  - Ensures compatibility in the number of functional and scalar predictors before computing the result.
```{r}
#' Inner product between two predictor_hybrid objects (with broadcasting)
#'
#' Computes the inner product between two `predictor_hybrid` objects,
#' including both functional and scalar components. Supports broadcasting
#' when one of the inputs has a single observation.
#'
#' @param xi_1 A `predictor_hybrid` object.
#' @param xi_2 Another `predictor_hybrid` object. If missing, defaults to `xi_1`.
#'
#' @return A numeric vector of inner products, or a scalar if both inputs contain a single observation.
#' @export
inprod.predictor_hybrid <- function(xi_1, xi_2 = NULL) {
  # Handle self-inner product
  if (is.null(xi_2)) xi_2 <- xi_1

  # Type checks
  if (!inherits(xi_1, "predictor_hybrid") || !inherits(xi_2, "predictor_hybrid")) {
    stop("Both inputs must be of class 'predictor_hybrid'.")
  }

  # Structural checks
  if (xi_1$n_functional != xi_2$n_functional) {
    stop("Mismatch in number of functional predictors.")
  }
  if (xi_1$n_scalar != xi_2$n_scalar) {
    stop("Mismatch in number of scalar predictors.")
  }

  # Swap so that broadcasting always applies to xi_2
  if (xi_1$n_sample == 1 && xi_2$n_sample > 1) {
    tmp <- xi_1
    xi_1 <- xi_2
    xi_2 <- tmp
  }

  n1 <- xi_1$n_sample
  n2 <- xi_2$n_sample

  if (!(n1 == n2 || n2 == 1)) {
    stop("Sample sizes are incompatible for broadcasting.")
  }

  # Prepare components
  f1 <- xi_1$functional_list
  f2 <- xi_2$functional_list
  Z1 <- xi_1$Z
  Z2 <- xi_2$Z

  # Replicate fd and Z if needed
  if (n2 == 1) {
    f2 <- rep_fd(f2, n1)
    Z2 <- matrix(rep(c(Z2), n1), nrow = n1, byrow = TRUE)
  }

  # Compute functional inner products
  inprod_functional <- vapply(seq_len(n1), function(i) {
    sum(vapply(seq_along(f1), function(j) {
      fda::inprod(f1[[j]][i], f2[[j]][i])
    }, numeric(1)))
  }, numeric(1))

  # Compute scalar inner products
  inprod_scalar <- rowSums(Z1 * Z2)

  # Combine results
  result <- inprod_functional + inprod_scalar
  if (n1 == 1 && n2 == 1) as.numeric(result) else result
}
```

```{r}
#' Inner product between two predictor_hybrid objects (with broadcasting)
#'
#' Computes the inner product between two `predictor_hybrid` objects,
#' including both functional and scalar components. Supports broadcasting
#' when one of the inputs has a single observation.
#'
#' @param xi_1 A `predictor_hybrid` object.
#' @param xi_2 Another `predictor_hybrid` object. If missing, defaults to `xi_1`.
#'
#' @return A numeric vector of inner products, or a scalar if both inputs contain a single observation.
#' @export
inprod_pen.predictor_hybrid <- function(xi_1, xi_2 = NULL, lambda) {

  inprod <- inprod.predictor_hybrid(xi_1, xi_2)
  # Prepare components

  f1 <- xi_1$functional_list

  f2 <- xi_2$functional_list
  for (j in 1:xi_1$n_functional){
    inprod <- inprod + lambda[j] * fda::inprod(
        fdobj1 = f1[[j]], 
        fdobj2 = f2[[j]], 
        Lfdobj1 = 2,
        Lfdobj2 = 2
        )

    }

  # Combine results
   return(as.numeric(inprod))
}
```

### subset_predictor_hybrid

```{r}
#' Extract a single observation from a predictor_hybrid object
#'
#' @param W A predictor_hybrid object with multiple samples.
#' @param i Integer index of the sample to extract.
#' @return A single-sample predictor_hybrid object.
subset_predictor_hybrid <- function(W, i) {
  new_Z <- matrix(W$Z[i, ], nrow = 1)
  new_functional_list <- lapply(W$functional_list, function(fdobj) {
    fd(coef = matrix(coef(fdobj)[, i], ncol = 1), basisobj = fdobj$basis)
  })
  new_predictor <- predictor_hybrid(Z = new_Z, functional_list = new_functional_list)
  return(new_predictor)
}
```

### replace_obs_hybrid

```{r}
#' Replace a single observation in a predictor_hybrid object
#'
#' Replaces the i-th observation of a predictor_hybrid object with a new
#' single-sample predictor_hybrid object.
#'
#' @param W A predictor_hybrid object with one or more samples.
#' @param i An integer index specifying the observation to replace.
#' @param new_W A single-sample predictor_hybrid object to use for replacement.
#'
#' @return A predictor_hybrid object with the i-th observation replaced.
#' @export
replace_obs_hybrid <- function(W, i, new_W) {
  # Input validation
  if (!inherits(W, "predictor_hybrid") || !inherits(new_W, "predictor_hybrid")) {
    stop("Both W and new_W must be of class 'predictor_hybrid'.")
  }
  if (new_W$n_sample != 1) {
    stop("new_W must be a single-sample predictor_hybrid object.")
  }
  if (i < 1 || i > W$n_sample) {
    stop(paste("Index i must be between 1 and", W$n_sample))
  }
  if (W$n_scalar != new_W$n_scalar) {
    stop("Mismatch in number of scalar predictors.")
  }
  if (W$n_functional != new_W$n_functional) {
    stop("Mismatch in number of functional predictors.")
  }

  # Check for compatible basis objects
  is_eqbasis <- getFromNamespace("is.eqbasis", "fda")
  for (j in seq_len(W$n_functional)) {
    if (!is_eqbasis(W$functional_list[[j]]$basis, new_W$functional_list[[j]]$basis)) {
      stop("Functional predictors must have the same basis objects.")
    }
  }

  # Replace the i-th observation in the scalar matrix Z
  W$Z[i, ] <- new_W$Z[1, ]

  # Replace the i-th observation in each functional predictor
  for (j in seq_along(W$functional_list)) {
    # The coefficients are stored as a matrix, with columns corresponding to samples
    W$functional_list[[j]]$coefs[, i] <- new_W$functional_list[[j]]$coefs[, 1]
  }

  # The number of samples remains the same
  return(W)
}
```

### add_broadcast(for matrix)

```{r}
#' @export
#' 
#' 
# --- 1. Define the S4 Generic Function ---
# This must be run first to tell R that 'add_broadcast' is a function 
# that can have methods defined for different classes.
setGeneric("add_broadcast", function(input, other, alpha = 1) {
  standardGeneric("add_broadcast")
})

# --- 2. Define the S4 Method for 'matrix' Class ---
setMethod("add_broadcast", "matrix",
###########################################
function(input, other, alpha = 1){

  # 1. Structural Check: Ensure RHS is a single observation.
  # This uses length() for vectors, and dim()[1] for matrices.
  is_matrix_other <- is.matrix(other)
  if (is_matrix_other && dim(other)[1] > 1) {
    stop("RHS must have a single observation (row).")
  }

  # 2. Conversion: Convert a single-row matrix into a vector for broadcasting.
  # If 'other' is already a vector, this does nothing.
  # If 'other' is a 1xP matrix, this flattens it into a length-P vector.
  if (is_matrix_other) {
    other <- as.vector(other) 
  }

  # 3. Native Arithmetic: Use simple R arithmetic for high performance.
  # R broadcasts the vector 'other' across the rows of the matrix 'input'.
  return( t( t(input) + alpha * other))
}
###########################################
)

```
### subtr_broadcast (for matrix)

```{r}
#' @export
#' 
#'
setGeneric("subtr_broadcast", function(input, other, alpha = 1) {
  standardGeneric("subtr_broadcast")
})
setMethod("subtr_broadcast", "matrix",
###########################################
function(input, other, alpha = 1){
  return(add_broadcast(input, other, (-1*alpha)))
  }
###########################################
)

```

# One iteration

## small functions

## penalty matrix construction
In this section, we provide an R function `get_constraint_matrix` that constructs the penalty matrix
$$J^*+\Lambda \ddot{J}^\ast,$$
as defined in Proposition 2 and used in the PLS component computation step of our proposed algorithm. We present functions  for computing $J^\ast$, $\Lambda$ and $\ddot{J}^\ast$ in sequence, and use these functions to define `get_constraint_matrix`.

### get_gram_matrix_block
Constructs a block-diagonal Gram matrix for a hybrid predictor object, defined as 
$$
    J^*=\mathrm{blkdiag}(J, I_p) \in \mathbb{R}^{(MK+p) \times (MK+p)}
$$
where
 $J = \mathrm{blkdiag}(J^{(1)}, \cdots, J^{(K)}) \in \mathbb{R}^{MK \times MK}$, and 
$$
J^{(k)} = \left[ \int_{\mathcal{T}k} b_m^{(k)}(t) , b_n^{(k)}(t) , dt \right]_{m,n=1}^M,
$$
as defined in \eqref{def:gram_basis} and \eqref{def:J_and_J_star}.

- Arguments
  - obj: A `predictor_hybrid` object containing both functional and scalar components.

- Value
  - A Matrix::bdiag sparse matrix representing the block-diagonal structure of the combined Gram matrix.

- Details: This function builds a block-diagonal matrix by:
  - Stacking the Gram matrices of each functional component (from obj$gram_list),
  - Appending an identity matrix of size equal to the number of scalar predictors (to represent unpenalized scalar covariates).
  The resulting matrix has size $(\texttt{total_dim} \times \texttt{total_dim})$, where total_dim = \sum_k M_k + p, with $M_k$ the number of basis functions for the $k$-th functional predictor and $p$ the number of scalar covariates.

- Usage
```r
block_gram <- get_gram_matrix_block(my_predictor)
```

**Code**
```{r}
#' Construct block-diagonal Gram matrix for hybrid predictor
#'
#' Returns a block-diagonal matrix containing the Gram  matrices for
#' each functional component and an identity matrix for the scalar part.
#'
#' @param obj A `predictor_hybrid` object.
#'
#' @return A block-diagonal matrix of size `(total_dim × total_dim)` where
#' functional and scalar components are arranged in order.
#' @export
get_gram_matrix_block <- function(obj) {
  if (!inherits(obj, "predictor_hybrid")) {
    stop("Input must be of class 'predictor_hybrid'.")
  }

  gram_blocks <- c(obj$gram_list, list(diag(obj$n_scalar)))
  Matrix::bdiag(gram_blocks)
}

```

**Unit test**
```{r}
testthat::test_that("get_gram_matrix_block constructs correct block-diagonal matrix", {
  suppressPackageStartupMessages(library(fda))
  suppressPackageStartupMessages(library(Matrix))

  # Construct hybrid predictor with two functional and three scalar predictors
  basis1 <- create.bspline.basis(c(0, 1), nbasis = 4)
  basis2 <- create.bspline.basis(c(0, 1), nbasis = 4)  # was 3; now valid

  fd1 <- fd(coef = matrix(1, 4, 2), basisobj = basis1)
  fd2 <- fd(coef = matrix(2, 4, 2), basisobj = basis2)
  Z <- matrix(rnorm(2 * 3), nrow = 2, ncol = 3)

  obj <- predictor_hybrid(Z = Z, functional_list = list(fd1, fd2))

  # Extract the block-diagonal Gram matrix
  G <- get_gram_matrix_block(obj)

  # Expected block sizes
  nb1 <- 4
  nb2 <- 4
  ns <- 3
  total_dim <- nb1 + nb2 + ns

  # Check matrix properties
  testthat::expect_equal(dim(G), c(total_dim, total_dim))

  # Check sub-block structure
  # Top-left: Gram matrix 1
  testthat::expect_equal(sum(G[1:nb1, 1:nb1]-obj$gram_list[[1]]), 0)

  # Next block: Gram matrix 2
  testthat::expect_equal(sum( G[(nb1 + 1):(nb1 + nb2), (nb1 + 1):(nb1 + nb2)]-obj$gram_list[[2]]),0 )

  # Final block: identity matrix for scalar part
  testthat::expect_equal( sum( G[(nb1 + nb2 + 1):total_dim, (nb1 + nb2 + 1):total_dim]  - diag(ns)), 0)
})

```
### get_smoothing_param_hybrid
Constructs a block-diagonal smoothing parameter matrix for use in penalized estimation involving hybrid predictors, denoted $\Lambda \in \mathbb{R}^{(MK+p) \times (MK+p)}$,  defined in \eqref{def:Lambda} as
$$
   \Lambda = \mathrm{blkdiag}(\lambda_1 I_M, \cdots, \lambda_K I_M, 0_{p \times p}) , 
$$

- Arguments
  - W: A predictor_hybrid object.
  - lambda: A numeric vector of length equal to the number of functional predictors. Each entry corresponds to a smoothing penalty weight.
- Value: A sparse block-diagonal matrix combining scaled identity matrices for the functional parts and a zero matrix for the scalar part.
- Details: This function generates a matrix used in regularized regression for functional predictors. Each functional component is penalized using a scaled identity matrix of its basis dimension. Scalar predictors are unpenalized.
- Usage:
```r
lambda_mat <- get_smoothing_param_hybrid(my_predictor, c(0.1, 0.2))
```
**code** 
```{r}
#' Construct block-diagonal smoothing parameter matrix
#'
#' Generates a block-diagonal matrix with smoothing parameters applied to each 
#' functional component. Each block is a scaled identity matrix, where the scaling 
#' factor corresponds to the regularization parameter for that functional component. 
#' The scalar components are not penalized and thus contribute a zero matrix block.
#'
#' @param W A `predictor_hybrid` object.
#' @param lambda A numeric vector of length equal to the number of functional components (`W$n_functional`), containing the smoothing parameters for each functional predictor.
#'
#' @return A block-diagonal matrix of size `(total_dim × total_dim)`, where the top-left blocks are scaled identity matrices for functional predictors and the bottom-right block is a zero matrix for scalar covariates.
#' @export
get_smoothing_param_hybrid <- function(W, lambda) {
  if (!inherits(W, "predictor_hybrid")) {
    stop("Input W must be of class 'predictor_hybrid'.")
  }

  if (length(lambda) != W$n_functional) {
    stop("Length of lambda must match the number of functional predictors.")
  }

  lambda_blocks <- lapply(seq_len(W$n_functional), function(ii) {
    nb <- W$functional_list[[ii]]$basis$nbasis
    lambda[ii] * diag(nb)
  })

  lambda_blocks[[W$n_functional + 1]] <- matrix(0, nrow = W$n_scalar, ncol = W$n_scalar)

  Matrix::bdiag(lambda_blocks)
}
```

**unit test**
```{r}
testthat::test_that("get_smoothing_param_hybrid returns correct block-diagonal structure", {
  suppressPackageStartupMessages(library(fda))
  basis1 <- create.bspline.basis(c(0, 1), nbasis = 5)
  basis2 <- create.bspline.basis(c(0, 1), nbasis = 4)

  fd1 <- fd(coef = matrix(1, 5, 3), basisobj = basis1)
  fd2 <- fd(coef = matrix(2, 4, 3), basisobj = basis2)

  Z <- matrix(1, nrow = 3, ncol = 2)

  obj <- predictor_hybrid(Z = Z, functional_list = list(fd1, fd2))
  lambda <- c(0.5, 2)

  lambda_mat <- get_smoothing_param_hybrid(obj, lambda)

  # Check overall dimension
  expected_dim <- sum(obj$n_basis_list) + obj$n_scalar
  testthat::expect_equal(dim(lambda_mat), c(expected_dim, expected_dim))

  # Check diagonal entries
  testthat::expect_equal(Matrix::diag(lambda_mat)[1:5], rep(0.5, 5))
  testthat::expect_equal(Matrix::diag(lambda_mat)[6:9], rep(2, 4))
  testthat::expect_equal(Matrix::diag(lambda_mat)[10:11], rep(0, 2))  # scalar part
})

```

### get_penalty_hybrid
Constructs a matrix defined in \label{def:J_dotdot_ast}
$$
    \ddot{J}^\ast=\mathrm{blkdiag}(\ddot{J}^{(1)}, \cdots, \ddot{J}^{(K)}, 0_{p \times p}) \in \mathbb{R}^{(MK+p) \times (MK+p)},
$$
where
 $\ddot{J}^{(k)}$ is  a $M \times M$ matrix defined in \label{def:dodot_J_k} as the gram matrix formed by the second derivativ of the basis functions
$$
\ddot{J}^{(k)} = \left[ \int_{\mathcal{T}_k} \ddot{b}^{(k)}_m(t) \ddot{b}^{(k)}_n(t) \, dt \right]_{m,n=1}^M.
$$

```{r}
testthat::test_that("get_penalty_hybrid computes correct dimensions and structure", {
  suppressPackageStartupMessages(library(fda))

  # Setup
  basis <- fda::create.bspline.basis(c(0, 1), nbasis = 5)
  n_sample <- 3
  n_scalar <- 2

  fd1 <- fda::fd(coef = matrix(1, 5, n_sample), basisobj = basis)
  fd2 <- fda::fd(coef = matrix(2, 5, n_sample), basisobj = basis)
  Z <- matrix(rnorm(n_sample * n_scalar), n_sample, n_scalar)
  obj <- predictor_hybrid(Z = Z, functional_list = list(fd1, fd2))

  # Run
  penalty <- get_penalty_hybrid(obj)

  # Expected dimensions
  total_dim <- sum(obj$n_basis_list) + obj$n_scalar
  testthat::expect_equal(dim(penalty), c(total_dim, total_dim))

  # Check block diagonal structure: bottom-right should be zero matrix
  scalar_block <- as.matrix(penalty[
    (total_dim - n_scalar + 1):total_dim,
    (total_dim - n_scalar + 1):total_dim
  ])
  testthat::expect_equal(scalar_block, matrix(0, n_scalar, n_scalar))

  # Check that the upper blocks are positive semidefinite
  eigenvalues <- eigen(as.matrix(penalty[1:(total_dim - n_scalar), 1:(total_dim - n_scalar)]))$values
  testthat::expect_true(all(eigenvalues >= -1e-8))  # Numerical tolerance
})

```

### get_constraint_matrix

```r
W <- predictor_hybrid(Z = Z, functional_list = list(fd1, fd2))
lambda <- c(0.1, 0.2)
constraint_matrix <- get_constraint_matrix_hybrid(W, lambda)
```

**code**
```{r}
#' Construct penalized constraint matrix for hybrid predictors
#'
#' Computes the denominator matrix used in penalized estimation for hybrid predictors,
#' combining the Gram matrix, smoothing parameter matrix, and penalty matrix.
#'
#' Specifically, this function returns the matrix:
#' \deqn{J^\ast + \Lambda \ddot{J}^\ast}
#' where \eqn{J^\ast} is the block-diagonal Gram matrix,
#' \eqn{\Lambda} is the block-diagonal smoothing parameter matrix,
#' and \eqn{\ddot{J}^\ast} is the block-diagonal penalty matrix of second derivative inner products.
#'
#' @param W A `predictor_hybrid` object.
#' @param lambda A numeric vector of smoothing parameters, one for each functional predictor.
#'
#' @return A matrix representing the penalized constraint matrix used in estimation.
#' @export
get_constraint_matrix_hybrid <- function(W, lambda) {
  J_star <- get_gram_matrix_block(W)
  J_dotdot_star <- get_penalty_hybrid(W)
  lambda_mat <- get_smoothing_param_hybrid(W, lambda)

  J_star + lambda_mat %*% J_dotdot_star
}

```

## PLS component computation

### Spectral method
The old implementation that computes the leading eigenvector of

### helper: get_CJ
```{r}
get_CJ <- function(W){
  n <- W$n_sample
  K <- W$n_functional
  ## C_J : C multiplied by J. Scalable to arbitrary K
  predictor_first <- W$functional_list[[1]]
  C_J <- t(predictor_first$coefs) %*% W$gram_list[[1]]
  for (i in 2:K){
    predictor_now <- W$functional_list[[i]]
    C_J <- cbind(
      C_J,
      t(predictor_now$coefs) %*% W$gram_list[[i]]
    )
    }
    return(C_J)
}
```

### helper: get_pre_corrcov 
construct the matrix V_star:
$$
    V^\ast = n^{-2}\begin{bmatrix}
         J \widetilde{C}^\top \mathbf{y}\mathbf{y}^\top \widetilde{C} J & J \widetilde{C}^\top \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{Z} \\
        \widetilde{Z}^\top \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{C} J &   \widetilde{Z}^\top \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{Z}
    \end{bmatrix} = n^{-2} J^* \widetilde{C}^{*\top} \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{C}^{*} J^*
    \in \mathbb{R}^{(MK+p) \times (MK+p)},
$$
where $J$ is defined in `get_gram_matrix_block`
equation (7), 
```{r}
#' Compute V matrix (pre-corrected covariance estimator)
#'
#' Constructs the V matrix involving inner products of functional predictors
#' (weighted by their Gram matrices) and scalar predictors against response y.
#'
#' @param W A `predictor_hybrid` object.
#' @param y A numeric response vector of length equal to number of samples.
#'
#' @return A matrix representing the corrected cross-covariance structure.
#' @export
get_pre_corrcov <- function(W, y){
  n <- W$n_sample
  K <- W$n_functional

  # building blocks
  ## C_J : C multiplied by J. Scalable to arbitrary K
  C_J <- get_CJ(W)

  ## covariance
  J_Ct_y <- t(C_J) %*% y
  Zt_y <- t(W$Z) %*% y

  # V matrix
  upper_left <- J_Ct_y %*% t(J_Ct_y)
  upper_right <- J_Ct_y %*% t(Zt_y)
  lower_left <- t(upper_right)
  lower_right <- Zt_y %*% t(Zt_y)

  V_star <- cbind(
    rbind(upper_left, lower_left),
    rbind(upper_right, lower_right)
    )/(n^2)

  return(V_star)
  }
```

### main function (wrapper: get_pls_comp
Computes the first PLS component from a hybrid predictor and response vector using regularization.

- Inputs
  - **`W`**: A `predictor_hybrid` object with functional and scalar predictors.
  - **`y`**: A numeric response vector of length equal to `W$n_sample`.
  - **`L`**: A Cholesky decomposition of a positive definite regularization matrix.
- Output: 
  - `xi_hat`: The first PLS component as a `predictor_hybrid` object
  - `E`: The matrix `inv(L) %*% V_star %*% t(inv(L))`
  - `V_star`: Cross-covariance matrix between predictors and response
  - `eigen_val`: Leading eigenvalue of `E`
```{r}
#' Compute the First PLS Component from a Hybrid Predictor
#'
#' Computes the coefficients of the first Partial Least Squares (PLS) component based on 
#' a hybrid predictor object and a response vector, using a regularized generalized eigenvalue problem.
#'
#' @param W A predictor_hybrid object containing both functional and scalar predictors.
#' @param y A numeric response vector of length equal to the number of samples in W.
#' @param L cholesky decompisition of a regularization matrix (typically positive definite) used in the generalized eigenproblem.
#'
#' @return A list with the following elements:
#'
#'   xi_hat: The estimated first PLS component as a predictor_hybrid object.
#'   E: The generalized eigenvalue problem matrix E = inv(L) V* t(inv(L)).
#'   V_star: The cross-covariance matrix between predictors and response.}
#'   eigen_val: The leading eigenvalue of E.
#' 
#'
#' @export
get_pls_comp <- function(W, y, L){

  V_star <- get_pre_corrcov(W, y)

  #invL <- Matrix::chol2inv(L)
  invL <- solve(L)
  A <- V_star %*% t(invL)       # A = V* × t(inv(L))
  E <- invL %*% A               # E = inv(L) × A

  eigen_result <- eigen(E)
  e <- eigen_result$vectors[, 1]
  if (is.complex(e)){
    print("stop")
    return("stop")
  }

  xi_star <- t(invL) %*% e      # xi_star solves t(L) xi = e
  xi_hat <- predictor_hybrid_from_coef(format = W, coef = xi_star)
  return(xi_hat)
}
```

## linear system method

### helper: get_xi_hat_linear
```{r}
get_xi_hat_linear_pen <- function(W, y, lambda){
  # new
  ## step 1: inner products
    n <- W$n_sample
  K <- W$n_functional
  u <- gamma <- list()

  v <- t(W$Z) %*% y
  q <- sum(v^2)
  for (j in 1:K){
    Theta_t <- W$functional_list[[j]]$coefs
    B <- W$gram_list[[j]]
    u[[j]] <- B %*% Theta_t %*% y

    # linear system
    R <- W$gram_list[[j]] + lambda[j] * W$gram_deriv2_list[[j]]
    gamma[[j]] <- solve(R, u[[j]])

     #q
    q <- q + sum( gamma[[j]] * ( R %*% gamma[[j]] ))
  }
  d_vec <- c(do.call(c, gamma), v)  # or: unlist(d)
  ## step 2: squared norm
  d_vec <- d_vec/ sqrt(q)
  xi_hat <- predictor_hybrid_from_coef(format = W, coef = d_vec)
  return(xi_hat)
}
```

the full vector of PLS scores $$\hat{\boldsymbol{\rho}}  = (\hat{\rho}_1, \ldots, \hat{\rho}_n)^\top$$
	is computed through the following matrix multiplication:
$$
		\hat{\boldsymbol{\rho}}^\top = \sum_{k=1}^K ( \hat{\boldsymbol{\gamma}}_j )^\top B \, \Theta_j^\top + \boldsymbol{\zeta}^\top \mathbf{Z}^\top.
$$	
```{r}
get_rho <- function(d_vec, W){
  K <- W$n_functional
  M <- W$n_basis_list[[1]]
  n_scalar <- W$n_scalar
  n <- W$n_sample
  rho <- rep(0, n)

  #functional part
  for (j in 1:K){
    start_index <- (j-1)*M + 1
    end_index <- j*M
    gamma_j <- d_vec[ start_index : end_index]
    Theta_j  <- t(W$functional_list[[j]]$coefs)
    B <- W$gram_list[[j]]
    rho <- rho + Theta_j %*% B %*% gamma_j
  }

  #scalar part
  rho <- rho + W$Z %*% d_vec[(K * M + 1):length(d_vec)]
  return (as.vector(rho))
}
```

```{r}
library(fda)

# --- Step 1: Simulate Data ---
set.seed(111)

n_sample <- 100
n_scalar <- 2
n_basis <- 20
n_functional <- 2

# Scalar predictors Z (centered)
Z <- matrix(rnorm(n_sample * n_scalar), ncol = n_scalar)
Z <- scale(Z, center = TRUE, scale = FALSE)

# Create functional predictors (centered)
basis <- create.bspline.basis(c(0, 1), nbasis = n_basis)
functional_list <- list()
for (j in 1:n_functional) {
  coefs <- matrix(rnorm(n_basis * n_sample), nrow = n_basis)
  fdobj <- fd(coef = coefs, basisobj = basis)
  mean_fd <- mean.fd(fdobj)
  mean_coef <- coef(mean_fd)
  centered_coefs <- coefs - matrix(mean_coef, nrow = n_basis, ncol = n_sample)
  functional_list[[j]] <- fd(coef = centered_coefs, basisobj = basis)
}

# --- Step 2: Construct hybrid object ---
W <- predictor_hybrid(Z = Z, functional_list = functional_list)

# --- Step 3: Simulate and center response ---
total_coef_len <- sum(W$n_basis_list) + W$n_scalar
true_coef <- rnorm(total_coef_len)

# Flatten design matrix
Xmat <- matrix(NA, n_sample, total_coef_len)
for (i in 1:n_sample) {
  W_i <- subset_predictor_hybrid(W, i)
  func_coefs <- do.call(c, lapply(W_i$functional_list, function(fdobj) as.vector(coef(fdobj))))
  Xmat[i, ] <- c(func_coefs, W_i$Z)
}

y <- Xmat %*% true_coef + rnorm(n_sample, sd = 0.1)
y <- as.vector(scale(y, center = TRUE, scale = FALSE))

# --- Step 4: Compute constraint matrix ---
lambda <- rep(0, W$n_functional)
#constr_mat <- get_constraint_matrix_hybrid(W, lambda)
#L <- chol(constr_mat)

# --- Step 5: Run eigen-based PLS ---
W_template <- W  # prevent mutation

pls_eig_result <- get_pls_comp(W, y, L)
xi_hat_eig <- pls_eig_result$xi_hat

xi_hat_pen <- get_xi_hat_linear_pen(W, y, lambda)
xi_hat_lin_new <- get_xi_hat_linear_new(W, y)
# Wrap into predictor_hybrid object using format template
xi_hat_direct <- predictor_hybrid_from_coef(format = W_template, coef = xi_hat_lin)
norm_val <- sqrt(inprod.predictor_hybrid(xi_hat_direct))
xi_hat_direct <- scalar_mul.predictor_hybrid(xi_hat_direct,  1 / norm_val)
xi_hat_direct_new <- predictor_hybrid_from_coef(format = W_template, coef = xi_hat_lin_new)

cat(xi_hat_direct_new$functional_list[[1]]$coefs- xi_hat_direct$functional_list[[1]]$coefs)
cat(xi_hat_direct_new$functional_list[[2]]$coefs- xi_hat_direct$functional_list[[2]]$coefs)

```

## Residualization

### response residualization

```{r}
get_nu <- function(y, rho){
            nu <- sum(y*rho) / sum(rho*rho)
            return(nu)  
}
```

```{r}
residualize_y <- function(y, rho, nu){
  y_next <- y - nu * rho
  return(y_next)
}
```

```{r}
LSE_ptws <- function(input, rho) {
  # old implementation
            rho_norm_sq <- (Matrix::norm(rho, type = "2"))^2
            nu <- as.numeric(t(rho) %*% input) / rho_norm_sq
            return(nu)
}

```

###  predictor residualization

$$
\widetilde{W}_i^{[l+1]} 
			:= 	\widetilde{W}_i^{[l]}  - 
			\widehat{\rho}_i^{[l]}
			\hat{\delta}^{[l]},~\text{where}~
			\delta^{[l]}
			:= 
			\frac{1}{\| 	\hat{\boldsymbol{\rho}}^{[l]}\|_2^2}
			\sum_{i=1}^n 
			\widehat{\rho}_i^{[l]}
			\widetilde{W}_i^{[l]},
$$

```{r}
get_delta <- function(W, rho){
  delta <- scalar_mul.predictor_hybrid(subset_predictor_hybrid(W, 1), rho[1])
   for (i in 2:length(rho)) {
  delta <- add.predictor_hybrid(delta, subset_predictor_hybrid(W, i),  rho[i]) 
   } 
  delta <- scalar_mul.predictor_hybrid(delta, 1/sum(rho*rho))
  return(delta)
}
```

```{r}
residualize_predictor <- function(W, rho, delta){
  n <- length(rho)
  W_res <- W
  for (i in 1:n){
    W_i <- subset_predictor_hybrid(W, i)
    W_res_i <-subtr.predictor_hybrid(W_i, delta,  rho[i]) 
    W_res <- replace_obs_hybrid(W_res, i, W_res_i)
  }
  return(W_res)
}
```

### LSE_hybrid

```{r}
LSE_hybrid <- function(W, rho) {
  # old implementation
  C_star_J_star <- matrix(NA, nrow = W$n_sample, ncol = sum(W$n_basis_list) + W$n_scalar)
  for (ii in (1 : W$n_functional)){
    cumsum_n_basis_now <- sum(W$n_basis_list[1:ii])
    n_basis_now <- W$n_basis_list[ii]
    C_star_J_star[, (cumsum_n_basis_now - n_basis_now + 1) :cumsum_n_basis_now] <-
      t(W$functional_list[[ii]]$coefs) %*% W$jacobian_list[[ii]]
  }
  C_star_J_star[ ,(sum(W$n_basis_list) + 1) : ncol(C_star_J_star)] <- W$Z
  rho_t_C_star_J_star <- t(rho) %*% C_star_J_star

  # matrices in the statement of Proposition 4
  J_star <-as.matrix(get_jacobian_hybrid(W))
  d_hat_star <- t( #basis coefficient of hybrid regression coefficientdelta
    solve(
      t((norm(rho, type="2"))^2 * J_star),
      t(rho_t_C_star_J_star)
      )
    )
  delta_hat <- hybrid_from_coef(W, d_hat_star) #hybrid regression coefficient
  return(delta_hat)
  }
```

# main algorithm

```{r eval=FALSE, include=FALSE}
#' @export
fit.hybridPLS <- function(W, y, n_iter, lambda) {
  # 1. Initialize storage
  W_now <- rho <- xi <- delta <- nu  <- iota <- beta <- list()

  # 3. Initialize residuals
  W_now[[1]] <- W
  y_now <- y
  for (l in 1:n_iter) {
    cat(paste(l, "th component", "\n"))

    xi[[l]]  <- get_xi_hat_linear_pen(W_now[[l]], y_now, lambda) # PLS direction
    rho[[l]] <- inprod.predictor_hybrid(W_now[[l]], xi[[l]]) #PLS score
    delta[[l]] <- get_delta(W_now[[l]], rho[[l]]);  
    W_now[[l + 1]] <- residualize_predictor(W_now[[l]], rho[[l]], delta[[l]]) # predictor residual
    nu[[l]] <- get_nu(y_now, rho[[l]]); y_now <- residualize_y(y_now, rho[[l]], nu[[l]]) # response residual

    iota[[l]] <- xi[[l]]
    if (l == 1) {
      beta[[l]] <- scalar_mul.predictor_hybrid( iota[[l]], nu[[l]] )
    } else {
      for (u in 1:(l - 1)) iota[[l]] <- subtr.predictor_hybrid( iota[[l]], iota[[u]], inprod.predictor_hybrid(delta[[u]], xi[[l]]))
      beta[[l]] <- add.predictor_hybrid(beta[[l - 1]], iota[[l]], nu[[l]])
    }
  }
  return (list(
    rho = rho,
    xi = xi,
    W = W_now,
    beta = beta
  ))
}

```

```{r eval=FALSE, include=FALSE}
#' @export
fit.hybridPLS_eigen <- function(W, y, n_iter, lambda) {
  # 1. Initialize storage
  W_now <- rho <- xi <- delta <- nu  <- sigma <- eta <- list()

  # 3. Initialize residuals
  W_now[[1]] <- W
  y_now <- y
  final_succesful_iteration <- 0
  constr_mat <- get_constraint_matrix_hybrid(W, lambda)
  constr_mat_chol <- t(chol(constr_mat))
  for (l in 1:n_iter) {
    cat(paste(l, "th component", "\n"))

    #xi[[l]]  <- get_xi_hat_linear_pen(W_now[[l]], y_now, lambda) # PLS direction
    xi[[l]]  <- get_pls_comp(W_now[[l]], y_now, constr_mat_chol) # PLS direction
    rho[[l]] <- inprod.predictor_hybrid(W_now[[l]], xi[[l]]) #PLS score
    delta[[l]] <- get_delta(W_now[[l]], rho[[l]]);  
    W_now[[l + 1]] <- residualize_predictor(W_now[[l]], rho[[l]], delta[[l]]) # predictor residual
    nu[[l]] <- get_nu(y_now, rho[[l]]); y_now <- residualize_y(y_now, rho[[l]], nu[[l]]) # response residual

    # Step 3: Orthogonalize and update prediction
    #sigma[[l]] <- xi[[l]]
    #if (l == 1) {
    #  eta[[l]] <- scalar_mul(sigma[[l]], nu[[l]])
    #} else {
    #  for (u in 1:(l - 1)) {
    #    sigma[[l]] <- subtr(sigma[[l]], sigma[[u]], hybrid_inner_prod(delta[[u]], xi[[l]]))
    #  }
    #  eta[[l]] <- add(eta[[l - 1]], sigma[[l]], nu[[l]])
    #}
  }
  return (list(
    rho = rho,
    xi = xi,
    W = W_now
  ))
}

```

```{r}
inprod.predictor_hybrid(res$W[[5]], res$xi[[3]])

sum(y * inprod.predictor_hybrid(res$W[[5]], res$xi[[3]]))

inprod_pen.predictor_hybrid(res$xi[[1]], res$xi[[8]], lambda) - inprod_pen.predictor_hybrid(res_eigen$xi[[1]], res_eigen$xi[[8]], lambda)
inprod_pen.predictor_hybrid(res$xi[[1]], res$xi[[8]], 10*lambda) - inprod_pen.predictor_hybrid(res_eigen$xi[[1]], res_eigen$xi[[8]], 10*lambda)
inprod_pen.predictor_hybrid(res$xi[[1]], res$xi[[8]], lambda/10) - inprod_pen.predictor_hybrid(res_eigen$xi[[1]], res_eigen$xi[[8]], lambda/10)
sum(y * inprod.predictor_hybrid(res$W[[8]], res$xi[[1]])) - sum(y * inprod.predictor_hybrid(res_eigen$W[[8]], res_eigen$xi[[1]]))

inprod_pen.predictor_hybrid(res$xi[[3]], res$xi[[4]], lambda) - inprod_pen.predictor_hybrid(res_eigen$xi[[3]], res_eigen$xi[[4]], lambda)
inprod_pen.predictor_hybrid(res$xi[[3]], res$xi[[4]], 10*lambda) - inprod_pen.predictor_hybrid(res_eigen$xi[[3]], res_eigen$xi[[4]], 10*lambda)
inprod_pen.predictor_hybrid(res$xi[[3]], res$xi[[4]], lambda/10)
sum(y * inprod.predictor_hybrid(res$W[[3]], res$xi[[1]]))

inprod_pen.predictor_hybrid(res$xi[[5]], res$xi[[10]], lambda)
inprod_pen.predictor_hybrid(res$xi[[5]], res$xi[[10]], 10*lambda)
inprod_pen.predictor_hybrid(res$xi[[5]], res$xi[[10]], lambda/10)
sum(y * inprod.predictor_hybrid(res$W[[5]], res$xi[[1]]))

inprod_pen.predictor_hybrid(res$xi[[4]], res$xi[[4]], lambda)
inprod_pen.predictor_hybrid(res$xi[[4]], res$xi[[4]], lambda/10)
inprod_pen.predictor_hybrid(res$xi[[4]], res$xi[[4]], 10*lambda )

```

```{r}
lambda <- c(0.5,0.5)
L=8
res <- fit.hybridPLS(W,y,L, lambda)
res_eigen <- fit.hybridPLS_eigen(W,y,L, lambda)
```

### nipals_pen_hybrid

- depends on:
  - 'get_constraint_matrix_hybrid'
  - 'get_pls_comp'
  - 'hybrid_inner_prod'
  - 'LSE_hybrid'

```{r eval=FALSE, include=FALSE}
#' Iterative Penalized NIPALS Algorithm for Hybrid PLS
#'
#' Performs iterative extraction of PLS components using a penalized NIPALS algorithm
#' for hybrid predictor objects that include both scalar and functional data.
#'
#' At each iteration:
#' \enumerate{
#'   \item Computes the first penalized PLS component using \code{get_pls_comp}.
#'   \item Computes PLS scores via hybrid inner products.
#'   \item Residualizes both predictors and response with respect to the extracted scores.
#'   \item Orthogonalizes and accumulates loading vectors to construct final linear predictors.
#' }
#' The same constraint matrix (computed once) is used across all iterations.
#'
#' @param W A \code{predictor_hybrid} object containing scalar and functional predictors.
#' @param y A numeric response vector of length equal to \code{W$n_sample}.
#' @param n_iter An integer specifying the number of PLS components (iterations) to compute.
#' @param lambda A numeric vector of smoothing parameters (length = number of functional predictors).
#' @param verbose Logical; if \code{TRUE}, prints progress for each component.
#'
#' @return This function operates by side effects. Internally, it constructs and updates:
#' \itemize{
#'   \item \code{xi}, \code{rho}, \code{delta}, \code{nu}: component vectors and regression effects,
#'   \item \code{sigma}, \code{eta}: orthogonalized components and accumulated prediction functions,
#'   \item \code{W_now}, \code{resid_y}: residualized predictor and response at each iteration,
#'   \item \code{E}, \code{V_star}, \code{eigen_val}: matrices for monitoring stability and eigenvalue progress.
#' }
#' The final number of successful iterations is tracked via \code{final_succesful_iteration}.
#'
#' @export
nipals_pen_hybrid <- function(W, y, n_iter, lambda, verbose) {
  # 1. Initialize storage
  rho <- xi <- delta <- nu  <- sigma <- eta <- list()
  E <- V_star <- eigen_val <- list()
  fitted_value_W <- fitted_value_y <- list()
  resid_y <- W_now <- list()
  first_eigen_val <- mse_W <- mse_y <- rep(NA, n_iter)

  # 2. Compute and store the constraint matrix once
  constr_mat <- get_constraint_matrix_hybrid(W, lambda)
  constr_mat_chol <- t(chol(constr_mat))

  # 3. Initialize residuals
  W_now[[1]] <- W
  y_now <- y
  final_succesful_iteration <- 0

  for (l in 1:n_iter) {
    if (verbose) cat(paste(l, "th component", "\n"))

    # Step 1: Compute penalized PLS direction and score
    pls_result_now <- get_pls_comp(W_now[[l]], y_now, constr_mat_chol, verbose)
    if (!is.list(pls_result_now)) break

    xi[[l]] <- pls_result_now$xi
    rho[[l]] <- hybrid_inner_prod(W_now[[l]], xi[[l]])

    # Step 2: Residualize predictor and response
    delta[[l]] <- LSE_hybrid(W_now[[l]], rho[[l]])
    fitted_value_W[[l]] <- fitted_value(delta[[l]], rho[[l]])
    W_now[[l + 1]] <- subtr(W_now[[l]], fitted_value_W[[l]])

    nu[[l]] <- LSE_ptws(y_now, rho[[l]])
    fitted_value_y[[l]] <- nu[[l]] * rho[[l]]
    y_now <- y_now - fitted_value_y[[l]]
    resid_y[[l]] <- y_now

    # Step 3: Orthogonalize and update prediction
    sigma[[l]] <- xi[[l]]
    if (l == 1) {
      eta[[l]] <- scalar_mul(sigma[[l]], nu[[l]])
    } else {
      for (u in 1:(l - 1)) {
        sigma[[l]] <- subtr(sigma[[l]], sigma[[u]], hybrid_inner_prod(delta[[u]], xi[[l]]))
      }
      eta[[l]] <- add(eta[[l - 1]], sigma[[l]], nu[[l]])
    }

    # Record diagnostics
    E[[l]] <- pls_result_now$E
    V_star[[l]] <- pls_result_now$V_star
    eigen_val[[l]] <- pls_result_now$eigen_val
    first_eigen_val[l] <- eigen_val[[l]][1]
    final_succesful_iteration <- l
  }
}

```

# Simulation tools

## sample splitting and normalization

```{r}
create_idx_train_test <-function(n_sample, train_ratio){
  n_train <- floor(n_sample*train_ratio)
  idx_train <- sort(sample(n_sample,n_train, replace=F))
  idx_test <- sort((1 : n_sample)[-idx_train])
  return(list(
    idx_train = idx_train, idx_test = idx_test
  ))
}
```

```{r}
#' @title Generate Train Indices
#' @description Generates train/test split indices.
#' @param n_sample Integer. Number of samples.
#' @param train_ratio Numeric. Ratio of train samples.
#' @return Vector of train indices.
get_idx_train <- function(n_sample, train_ratio) {
  n_train <- floor(n_sample * train_ratio)
  idx_train <- sort(sample(n_sample, n_train, replace = FALSE))
  return(idx_train)
}
```

```{r}
create_idx_kfold <- function(n_sample, n_fold){
  idx_list <- caret::createFolds( (1 : n_sample), k = n_fold, list = TRUE, returnTrain = FALSE)
  idx_list_list <- list()
  for (i in 1:length(idx_list)){
    valid_hyper_index <- idx_list[[i]]
    train_hyper_index <- (1 : n_sample)[-valid_hyper_index]
    idx_list_list[[i]] <- list("idx_train" = train_hyper_index, "idx_valid" = valid_hyper_index)
  }
  return(idx_list_list)
}
```

```{r}
n_sample.fd <- function(fd_obj){
  length(fd_obj$fdnames$reps)
}
```

### split.all

Splits functional predictors, scalar predictors, and response into train and test sets.

```{r}
#' Split Hybrid Predictor and Response Data
#'
#' Splits the hybrid predictor object and the response vector into training and
#' testing sets based on a given training ratio.
#'
#' @param W_hybrid predictor_hybrid object containing all functional and scalar data.
#' @param response Vector. The response variable corresponding to the samples in W_hybrid.
#' @param train_ratio Numeric scalar between 0 and 1 indicating the proportion of data to use for training.
#' @return List containing split train and test sets for all components.
#' @export
split.all <- function(W_hybrid, response, train_ratio) {
  # --- Type and Ratio Checks ---
  stopifnot(
    "'W_hybrid' must be a predictor_hybrid object" = inherits(W_hybrid, "predictor_hybrid"),
    "'response' must be a vector" = is.vector(response),
    "Response length must match hybrid sample size" = length(response) == W_hybrid$n_sample,
    "'train_ratio' must be between 0 and 1" = is.numeric(train_ratio) && train_ratio > 0 && train_ratio < 1
  )

  N <- W_hybrid$n_sample
  N_train <- floor(N * train_ratio)

  # Generate random indices for training set
  # Using 1:N ensures indices are always valid
  train_idx <- sample(1:N, N_train)
  test_idx <- (1:N)[-train_idx]

  W_train <- predictor_hybrid(
    W_hybrid$Z[train_idx, , drop = FALSE],
    list(
      W_hybrid$functional_list[[1]][train_idx],
      W_hybrid$functional_list[[2]][train_idx]
    )
  )

  W_test <- predictor_hybrid(
    W_hybrid$Z[test_idx, , drop = FALSE],
    list(
      data_fun_1_test = W_hybrid$functional_list[[1]][test_idx],
    data_fun_2_test = W_hybrid$functional_list[[2]][test_idx]
    )
  )

  return( 
    list(
    predictor_train = W_train,
    predictor_test = W_test ,
    response_train = response[train_idx],
    response_test = response[test_idx]
  ))
}
```

### curve_normalize
Subtracts the mean function and divides by a scaling constant (deno). This implementation does not use the native fda S3 methods (minus.fd and times.fd), since they do not broadcast well.
```{r}
#' 
#' @export
curve_normalize <- function(functional, mean_functional, deno){
  # operation also takes care of the original_X

  functional_normalized <- fd(coef =
                                coef(functional) - coef(mean_functional) %*% matrix(1, ncol = ncol(coef(functional))),
                              basisobj = functional$basis
                              )#numerator
  functional_normalized <- times.fd(1/deno, functional_normalized)
  return(functional_normalized)
}
```

### curve_normalize_train_test
Normalizes the functional components of both the training and testing 'predictor_hybrid' objects using the mean and standard deviation derived ONLY from the training data. The scalar components (Z) are left unchanged.

```{r}
#' Normalize Functional Predictors in Hybrid Objects (Train/Test Split)
#' @param train A `predictor_hybrid` object representing the training set.
#' @param test A `predictor_hybrid` object representing the testing set.
#' @return A list containing:
#'   train: The normalized training `predictor_hybrid` object.}
#'   test: The normalized testing `predictor_hybrid` object.}
#'   deno_train: A numeric vector of scaling factors used for each functional predictor.}
#' @export
curve_normalize_train_test <- function(train, test) {
  # --- Type Checks ---
  stopifnot(
    "'train' must be of class 'predictor_hybrid'" = inherits(train, "predictor_hybrid"),
    "'test' must be of class 'predictor_hybrid'" = inherits(test, "predictor_hybrid"),
    "Mismatch in number of functional predictors" = train$n_functional == test$n_functional
  )

  train_normalized <- train
  test_normalized <- test
  deno_train <- numeric(train$n_functional)
  mean_train <- list()

  # Iterate through each functional predictor (k)
  for (k in seq_len(train$n_functional)) { 
    # 1. Determine centering and scaling factors from the TRAINING data only
    train_fd <- train$functional_list[[k]]

    # Calculate the mean function for centering
    mean_train[[k]] <- fda::mean.fd(train_fd)

    # Calculate the standard deviation function
    sd_train <- fda::sd.fd(train_fd)

    # Calculate the scaling factor (L2 norm of the standard deviation function)
    deno_train[k] <- sqrt(fda::inprod(sd_train, sd_train))

    # Ensure scaling factor is positive
    if (deno_train[k] <= 1e-10) {
      warning(paste("Scaling factor for functional predictor", k, "is near zero. Skipping normalization for this predictor."))
      deno_train[k] <- 1 # Set to 1 to prevent division by zero, effectively no normalization
    }

    # 2. Normalize the TRAINING set
    train_normalized$functional_list[[k]] <- curve_normalize(
      train_fd, # Use original train data
      mean_train[[k]],
      deno_train[k]
    )

    # 3. Normalize the TEST set (using train factors to prevent leakage)
    test_normalized$functional_list[[k]] <-
      curve_normalize(
        test$functional_list[[k]], # Use original test data
        mean_train[[k]], # Centered by training mean
        deno_train[k] # Scaled by training SD
      )
  }
  # Return the modified hybrid objects and the scaling factors
  return( list(
    predictor_train = train_normalized,
    predictor_test = test_normalized,
    mean_train = mean_train,
    deno_train = deno_train
  ))
}
```

```{r}
#'
#' @export
scalar_normalize <- function(scalar_predictor, mean, deno){
   scalar_predictor_normalized <- subtr_broadcast(scalar_predictor, mean) #numerator
   deno_mat <- matrix(rep(1,nrow(scalar_predictor))) %*% deno
   scalar_predictor_normalized <- scalar_predictor_normalized / deno_mat
   return(scalar_predictor_normalized)
 }
```

```{r}
#' Normalize the scalar component (Z) of hybrid predictors using training set statistics.
#'
#' @param predictor_train A `predictor_hybrid` object (train set).
#' @param predictor_test A `predictor_hybrid` object (test set).
#' @return A list containing the normalized 'train' and 'test' hybrid objects.
#' @export
scalar_normalize_train_test <- function(predictor_train, predictor_test){

  # Type and structure checks
  stopifnot(
    inherits(predictor_train, "predictor_hybrid"),
    inherits(predictor_test, "predictor_hybrid"),
    predictor_train$n_scalar == predictor_test$n_scalar
  )

  # Create copies for modification
  predictor_train_normalized <- predictor_train
  predictor_test_normalized <- predictor_test

  # 1. Calculate statistics from the TRAINING set only
  mean_train <- colMeans(predictor_train$Z)
  sd_train <- apply(predictor_train$Z, 2, sd)

  # 2. Apply normalization to the TRAINING set
  predictor_train_normalized$Z <-
    scalar_normalize(
      predictor_train$Z,
      mean_train,
      sd_train
    )

  # 3. Apply normalization to the TEST set using TRAIN statistics
  predictor_test_normalized$Z <-
    scalar_normalize(
      predictor_test$Z,
      mean_train, # IMPORTANT: Use train mean
      sd_train    # IMPORTANT: Use train SD
    )

  return(
    list(
      predictor_train = predictor_train_normalized,
      predictor_test = predictor_test_normalized,
      mean_train = mean_train,
      sd_train = sd_train
    )
  )
}
```

```{r}
#' Scale the scalar predictors so that the variability between the
#' functional and scalar predictors are comparable.
#'
#' This method calculates a scaling factor (sqrt(omega)) from the training
#' set, defined as the ratio of the total sum of squares (L2 norms) of all
#' functional components to the total sum of squares of all scalar components.
#'
#' @param train A predictor_hybrid object representing the training data.
#' @param test A predictor_hybrid object representing the test data.
#'
#' @return A list containing the normalized 'train' and 'test' predictor_hybrid objects.
#' @export
btwn_normalize_train_test <- function(train, test){

  # Type checks
  stopifnot(
    "'train' must be of class 'predictor_hybrid'" = inherits(train, "predictor_hybrid"),
    "'test' must be of class 'predictor_hybrid'" = inherits(test, "predictor_hybrid")
  )

  # Create copies for modification
  train_normalized <- train
  test_normalized <- test

  # Z_s for clarity in calculation
  Z_s <- train$Z

  # --- 1. Calculate the numerator of omega (Sum of L2 norms of all functional samples in train) ---
  omega_numerator <- 0

  for (k in seq_len(train$n_functional)){ # for kth functional predictor
    # fda::inprod(fd, fd) returns the N x N matrix of inner products.
    # The diagonal elements are the L2 norm squared for each sample.
    functional_inprod_matrix <- fda::inprod(
      train$functional_list[[k]],
      train$functional_list[[k]]
    )

    # Sum the diagonal to get the total squared norm for this predictor k
    omega_numerator <- omega_numerator + sum(diag(functional_inprod_matrix))
  }

  # --- 2. Calculate the denominator of omega (Sum of squares of all scalar samples in train) ---
  omega_denominator <- sum(Z_s^2)

  # Handle case where scalar variance is zero (prevent division by zero)
  if (omega_denominator == 0) {
    warning("Scalar component sum of squares is zero. No scaling applied.")
    omega <- 1
  } else {
    # Calculate the scaling factor omega
    omega <- omega_numerator / omega_denominator
  }

  # --- 3. Apply scaling factor sqrt(omega) to both train and test scalar components ---
  scaling_factor <- sqrt(omega)

  train_normalized$Z <- scaling_factor * (train$Z)
  test_normalized$Z <- scaling_factor * (test$Z)

  return(
    list(
      predictor_train = train_normalized,
      predictor_test = test_normalized,
      scaling_factor = scaling_factor
    )
  )
}

```

### normailize.all

```{r}
#' 
#' @export
split_and_normailize.all <- function(W_hybrid, response, train_ratio){

# $W_hybrid =  W_hybrid_mock
#$ response

# train_ratio = TRAIN_RATIO

 # split.result <- split.all(W_hybrid, response, train_ratio)

 #     predictor_train = W_train,
 #   predictor_test = W_test ,
 #   response_train = response[train_idx],
 #   response_test = response[test_idx]

 # list(
 #   train = train_normalized,
 #   test = test_normalized,
 #   mean_train = mean_train,
 #   deno_train = deno_train
#  ))

  curve_normalize_result <- curve_normalize_train_test(split.result$predictor_train, split.result$predictor_test)
  scalar_normalize_result <- scalar_normalize_train_test(curve_normalize_result$predictor_train, curve_normalize_result$predictor_test)
  btwn_normalize_result <- btwn_normalize_train_test(
    scalar_normalize_result$predictor_train,
    scalar_normalize_result$predictor_test
  )

  response_mean_train <-  mean(split.result$response_train)
  response_sd_train <-  sd(split.result$response_train)
  response_train <- (split.result$response_train - response_mean_train)/response_sd_train
  response_test <- (split.result$response_test - response_mean_train)/response_sd_train
  response_normalize_result = list(mean_train = response_mean_train, sd_train = response_sd_train)
  return(
    list(
      predictor_train = btwn_normalize_result$predictor_train,
      predictor_test = btwn_normalize_result$predictor_test,
      response_train = response_train,
      response_test = response_test,
      curve_normalize_result = curve_normalize_result,
      scalar_normalize_result = scalar_normalize_result,
      btwn_normalize_result = btwn_normalize_result,
      response_normalize_result = response_normalize_result
    )
  )
}
```

# Baseline method: FPCA regression

## Modified MFPCA Function for Simultaneous Training and Prediction

The MFPCA (Multivariate Functional Principal Component Analysis) function computes multivariate functional PCA for functions defined across different (dimensional) domains.
We have modified the original implementation to admit a new input, **`mFData_predict`**, allowing the function to calculate PCA scores for **both the training data and new test (or validation) data simultaneously** in a single run. This change was necessary because the standard `predict.MFPCAfit` function does not permit computing scores for new, out-of-sample functions.
The multivariate functional principal component analysis inherently relies on a **univariate basis expansion** for each functional element ($X^{(j)}$). To simplify the code modification, we use only the univariate functional PCA option and have removed other options provided by the original MFPCA function. We modifies the main function and two internal functions.

### MFPCA

```{r}
#'
#' @export MFPCA
MFPCA <- function(mFData, mFData_predict, M, uniExpansions, weights = rep(1, length(mFData)), fit = FALSE, approx.eigen = FALSE,
                  bootstrap = FALSE, nBootstrap = NULL, bootstrapAlpha = 0.05, bootstrapStrat = NULL,
                  verbose = options()$verbose)
{
  if(! inherits(mFData, "multiFunData"))
    stop("Parameter 'mFData' must be passed as a multiFunData object.")

  # number of components
  p <- length(mFData)
  # number of observations
  N <- nObs(mFData)

  if(!all(is.numeric(M), length(M) == 1, M > 0))
    stop("Parameter 'M' must be passed as a number > 0.")

  if(!(is.list(uniExpansions) & length(uniExpansions) == p))
    stop("Parameter 'uniExpansions' must be passed as a list with the same length as 'mFData'.")

  if(!(is.numeric(weights) & length(weights) == p))
    stop("Parameter 'weights' must be passed as a vector with the same length as 'mFData'.")

  if(!is.logical(fit))
    stop("Parameter 'fit' must be passed as a logical.")

  if(!is.logical(approx.eigen))
    stop("Parameter 'approx.eigen' must be passed as a logical.")

  if(!is.logical(bootstrap))
    stop("Parameter 'bootstrap' must be passed as a logical.")

  if(bootstrap)
  {
    if(is.null(nBootstrap))
      stop("Specify number of bootstrap iterations.")

    if(any(!(0 < bootstrapAlpha & bootstrapAlpha < 1)))
      stop("Significance level for bootstrap confidence bands must be in (0,1).")

    if(!is.null(bootstrapStrat))
    {
      if(!is.factor(bootstrapStrat))
        stop("bootstrapStrat must be either NULL or a factor.")

      if(length(bootstrapStrat) != nObs(mFData))
        stop("bootstrapStrat must have the same length as the number of observations in the mFData object.")
    }
  }

  if(!is.logical(verbose))
    stop("Parameter 'verbose' must be passed as a logical.")

  # dimension for each component
  dimSupp <- dimSupp(mFData)

  # get type of univariate expansions
  type <- vapply(uniExpansions, function(l){l$type}, FUN.VALUE = "")

  # de-mean functions -> coefficients are also de-meaned!
  # do not de-mean in uFPCA, as PACE gives a smooth estimate of the mean (see below)
  m <- meanFunction(mFData, na.rm = TRUE) # ignore NAs in data
  for(j in seq_len(p))
  {
    if(type[j] != "uFPCA")
      mFData[[j]] <- mFData[[j]] - m[[j]]
  }

  if(verbose)
    cat("Calculating univariate basis expansions (", format(Sys.time(), "%T"), ")\n", sep = "")

########### Below is modified by Jongmin Mun #############################
  # calculate univariate basis expansion for all components
  uniBasis <- mapply(
    function(expansion, data, data_predict){
      do.call(univDecomp, c(
        list(funDataObject = data, funDataObject_predict = data_predict),
        expansion)
        )
      },
    expansion = uniExpansions,
    data = mFData,
    data_predict = mFData_predict,
    SIMPLIFY = FALSE
    )
########### Above is modified by Jongmin Mun #############################

  # for uFPCA: replace estimated mean in m
  for(j in seq_len(p))
  {
    if(type[j] == "uFPCA")
      m[[j]] <- uniBasis[[j]]$meanFunction
  }

  # Multivariate FPCA
  npc <- vapply(uniBasis, function(x){dim(x$scores)[2]}, FUN.VALUE = 0) # get number of univariate basis functions

  if(M > sum(npc))
  {
    M <- sum(npc)
    warning("Function MFPCA: total number of univariate basis functions is smaller than given M. M was set to ", sum(npc), ".")
  }

  # check if non-orthonormal basis functions used
  if(all(foreach::foreach(j = seq_len(p), .combine = "c")%do%{uniBasis[[j]]$ortho}))
    Bchol = NULL
  else
  {
    # Cholesky decomposition of B = block diagonal of Cholesky decompositions
    Bchol <- Matrix::bdiag(lapply(uniBasis, function(l){
      if(l$ortho)
        res <- Matrix::Diagonal(n = ncol(l$scores))
      else
        res <- Matrix::chol(l$B)

      return(res)}))
  }

  if(verbose)
    cat("Calculating MFPCA (", format(Sys.time(), "%T"), ")\n", sep = "")

  mArgvals <- if (utils::packageVersion("funData") <= "1.2") {
    getArgvals(mFData)
  } else {
    funData::argvals(mFData)
  }

  res <- calcMFPCA(N = N, p = p, Bchol = Bchol, M = M, type = type, weights = weights,
                   npc = npc, argvals = mArgvals, uniBasis = uniBasis, fit = fit, approx.eigen = approx.eigen)

  res$meanFunction <- m # return mean function, too

  names(res$functions) <- names(mFData)

  #############modified by jongmin mun

  #####################################
  if(fit)
  {
    res$fit <- m + res$fit # add mean function to fits
    names(res$fit) <- names(mFData)
  }

  # give correct names
  namesList <- lapply(mFData, names)
  if(!all(vapply(namesList, FUN = is.null, FUN.VALUE = TRUE)))
  {
    if(length(unique(namesList)) != 1)
      warning("Elements have different curve names. Use names of the first element for the results.")

    row.names(res$scores) <- namesList[[1]]

    if(fit)
      for(i in seq_len(p))
        names(res$fit[[i]]) <- namesList[[1]]
  }

  if(type[1] == "uFPCA"){
    scores.pred<-calcMFPCA_predict(N = nObs(mFData_predict),
                      p = p,
                      M = M,
                      weights = weights,
                      npc = npc,
                      uniBasis = uniBasis,
                      normFactors = res$normFactors,
                      vectors_train=res$vectors,
                      values_train=res$values
    )
  }
  res$scores.pred <- scores.pred

  class(res) <- "MFPCAfit"
    return(res)
}

```

### fpcaBasis

```{r}
################# modified by Jongmin Mun #####################
fpcaBasis <- function(funDataObject, funDataObject_predict, nbasis = 10, pve = 0.99, npc = NULL, makePD = FALSE, cov.weight.type = "none")
{
  FPCA <- PACE(funDataObject, predData = NULL, nbasis, pve, npc, makePD, cov.weight.type)
  FPCA_pred <- PACE(funDataObject = funDataObject, predData = funDataObject_predict, nbasis, pve, npc, makePD, cov.weight.type)
  return(list(scores = FPCA$scores,
              scores.pred = FPCA_pred$scores,
              ortho = TRUE,
              functions = FPCA$functions,
              meanFunction = FPCA$mu
  ))
}
```

### univDecomp
```{r}
univDecomp <- function(type, funDataObject, funDataObject_predict, ...)
{
  # Parameter checking
  if(is.null(type))
    stop("Parameter 'type' is missing.")
  else
  {
    if(!is.character(type))
      stop("Parameter 'type' must be a character string. See ?univDecomp for details.")
  }

  if(class(funDataObject) != "funData")
    stop("Parameter 'funDataObject' must be a funData object.")

  # get all arguments (except for function call and type)
  params <- list(...)

  # check if type and data are of correct type
  if(is.null(type))
    stop("univDecomp: must specify 'type'.")

  if(!inherits(type, "character"))
    stop("univDecomp: 'type' must be of class character.")

  if(is.null(funDataObject))
    stop("univDecomp: must specify 'funDataObject'.")

  if(class(funDataObject) != "funData")
    stop("univDecomp: 'funDataObject' must be of class funData.")

  params$funDataObject <- funDataObject # add funDataObject (-> make sure is evaluated in correct env.)

  ############### modified by Jongmin Mun ######################
  if (type == "uFPCA"){params$funDataObject_predict <- funDataObject_predict}
  ############### modified by Jongmin Mun ######################

  res <- switch(type,
                "given" = do.call(givenBasis, params),
                "uFPCA" = do.call(fpcaBasis, params),
                "UMPCA" = do.call(umpcaBasis, params),
                "FCP_TPA" = do.call(fcptpaBasis, params),
                "splines1D" = do.call(splineBasis1D, params),
                "splines1Dpen" = do.call(splineBasis1Dpen, params),
                "splines2D" = do.call(splineBasis2D, params),
                "splines2Dpen" = do.call(splineBasis2Dpen, params),
                "fda" = do.call(fdaBasis, params),
                "DCT2D" = do.call(dctBasis2D, params),
                "DCT3D" = do.call(dctBasis3D, params),
                stop("Univariate Decomposition for 'type' = ", type, " not defined!")
  )

  if(res$ortho == FALSE & is.null(res$B))
    stop("UnivDecomp: must provide integral matrix B for non-orthonormal basis functions.")

  return(res)
}
```

## Unmodified functions

### calcBasisIntegrals
```{r}

# define global variable j, used by the foreach package and confusing R CMD CHECK
globalVariables('j')

#' Utility function that calculates matrix of basis-scalar products (one dimension)
#'
#' If the element \eqn{X^{(j)}}{X^{(j)}} is expanded in basis functions \eqn{b_i^{(j)}(t),~ i = 1, \ldots, K_j}{b_i(t)},
#' this function calculates the \eqn{K_j \times K_j}{K_j  \times K_j} matrix \eqn{B^{(jj)}}{B^{(jj)}} with entries
#' \deqn{B^{(jj)}_{mn} = \int_{\mathcal{T_j}} b_m^{(j)}(t) b_n^{(j)}(t) \mathrm{d} t}.
#'
#' @section Warning: This function is implemented only for functions on one- or two-dimensional domains.
#'
#' @param basisFunctions Array of \code{npc} basis functions of dimensions \code{npc x M1} or \code{npc x M1 x M2}.
#' @param dimSupp dimension of the support of the basis functions (1 or 2)
#' @param argvals List of corresponding x-values.
#'
#' @return A matrix containing the scalar product of all combinations of basis functions (matrix \eqn{B^{(j)}})
#'
#' @seealso \code{\link{MFPCA}}, \code{\link[funData]{dimSupp}}
#'
#' @keywords internal
calcBasisIntegrals <- function(basisFunctions, dimSupp, argvals)
{
  npc <- dim(basisFunctions)[1]

  #  integral basis matrix
  B <- array(0, dim = c(npc, npc))

  if(dimSupp == 1) # one-dimensional domain
  {
    w <- funData::.intWeights(argvals[[1]])

    for(m in seq_len(npc))
    {
      for(n in seq_len(m))
        B[m, n] <- B[n, m] <- (basisFunctions[m, ]* basisFunctions[n, ])%*%w
    }
  }
  else # two-dimesional domain (otherwise function stops before!)
  {
    w1 <- t(funData::.intWeights(argvals[[1]]))
    w2 <- funData::.intWeights(argvals[[2]])

    for(m in seq_len(npc))
    {
      for(n in seq_len(m))
        B[m, n] <- B[n, m] <-  w1 %*%(basisFunctions[m, , ]* basisFunctions[n, ,])%*%w2
    }
  }

  return(B)
}
```

### calcMFPCA
```{r}
#' Internal function that implements the MFPCA algorithm for given univariate decompositions
#' @export
calcMFPCA <- function(N, p, Bchol, M, type, weights, npc, argvals, uniBasis, fit = FALSE, approx.eigen = FALSE)
{
  # combine all scores
  allScores <- foreach::foreach(j = seq_len(p), .combine = "cbind")%do%{uniBasis[[j]]$scores}

  # block vector of weights
  allWeights <- foreach::foreach(j = seq_len(p), .combine = "c")%do%{rep(sqrt(weights[j]), npc[j])}

  Z <- allScores %*% Matrix::Diagonal(x = allWeights) / sqrt(N-1)

  # check if approximation is appropriate (cf. irlba)
  if(approx.eigen & (M > min(N, sum(npc))/2))
  {
    warning("Calculating a large percentage of principal components, approximation may not be appropriate.
            'approx.eigen' set to FALSE.")
    approx.eigen = FALSE
  }

  # check if non-orthonormal basis functions used and calculate PCA on scores
  if(is.null(Bchol))
  {
    if(approx.eigen)
    {
      tmpSVD <- irlba::irlba(as.matrix(Z), nv = M)

      vectors <- tmpSVD$v
      values <- tmpSVD$d[seq_len(M)]^2
    }
    else
    {
      if(sum(npc) > 1000)
        warning("MFPCA with > 1000 univariate eigenfunctions and approx.eigen = FALSE. This may take some time...")

      e <- eigen(stats::cov(allScores) * outer(allWeights, allWeights, "*"))

      values <- e$values[seq_len(M)]
      vectors <- e$vectors[,seq_len(M)]
    }
  }
  else
  {
    if(approx.eigen)
    {
      tmpSVD <- irlba::irlba(as.matrix(Matrix::tcrossprod(Z, Bchol)), nv = M)

      vectors <- Matrix::crossprod(Bchol, tmpSVD$v)
      values <- tmpSVD$d[seq_len(M)]^2
    }
    else
    {
      if(sum(npc) > 1000)
        warning("MFPCA with > 1000 univariate eigenfunctions and approx.eigen = FALSE. This may take some time...")

      e <- eigen(Matrix::crossprod(Bchol) %*% (stats::cov(allScores) * outer(allWeights, allWeights, "*")))

      values <- Re(e$values[seq_len(M)])
      vectors <- Re(e$vectors[,seq_len(M)])
    }
  }

  # normalization factors
  normFactors <- 1/sqrt(diag(as.matrix(Matrix::crossprod(Z %*% vectors))))

  ### Calculate scores
  scores <- Z %*% vectors * sqrt(N-1) # see defintion of Z above!
  scores <- as.matrix(scores %*% diag(sqrt(values) * normFactors, nrow = M, ncol = M)) # normalization

  ### Calculate eigenfunctions (incl. normalization)
  npcCum <- cumsum(c(0, npc)) # indices for blocks (-1)

  tmpWeights <- as.matrix(Matrix::crossprod(Z, Z %*%vectors))
  eFunctions <- foreach::foreach(j = seq_len(p)) %do% {
    univExpansion(type = type[j],
                  scores = 1/sqrt(weights[j] * values) * normFactors * t(tmpWeights[npcCum[j]+seq_len(npc[j]), , drop = FALSE]),
                  argvals = argvals[[j]],
                  functions = uniBasis[[j]]$functions,
                  params = uniBasis[[j]]$settings)
  }

  res <- list(values = values,
              functions = multiFunData(eFunctions),
              scores = scores,
              vectors = vectors,
              values = values,
              normFactors = normFactors,
              uniBasis = uniBasis,
              uniExpansions = uniExpansions
  )

  # calculate truncated Karhunen-Loeve representation (no mean here)
  if(fit)
    res$fit <- multivExpansion(multiFuns = res$functions, scores = scores)

  return(res)
}
```

### calcMFPCA

```{r}
#' Internal function that implements the MFPCA algorithm for given univariate decompositions
#' @export
#' 
calcMFPCA <- function(N, p, Bchol, M, type, weights, npc, argvals, uniBasis, fit = FALSE, approx.eigen = FALSE)
{
  # combine all scores
  allScores <- foreach::foreach(j = seq_len(p), .combine = "cbind")%do%{uniBasis[[j]]$scores}

  # block vector of weights
  allWeights <- foreach::foreach(j = seq_len(p), .combine = "c")%do%{rep(sqrt(weights[j]), npc[j])}

  Z <- allScores %*% Matrix::Diagonal(x = allWeights) / sqrt(N-1)

  # check if approximation is appropriate (cf. irlba)
  if(approx.eigen & (M > min(N, sum(npc))/2))
  {
    warning("Calculating a large percentage of principal components, approximation may not be appropriate.
            'approx.eigen' set to FALSE.")
    approx.eigen = FALSE
  }

  # check if non-orthonormal basis functions used and calculate PCA on scores
  if(is.null(Bchol))
  {
    if(approx.eigen)
    {
      tmpSVD <- irlba::irlba(as.matrix(Z), nv = M)

      vectors <- tmpSVD$v
      values <- tmpSVD$d[seq_len(M)]^2
    }
    else
    {
      if(sum(npc) > 1000)
        warning("MFPCA with > 1000 univariate eigenfunctions and approx.eigen = FALSE. This may take some time...")

      e <- eigen(stats::cov(allScores) * outer(allWeights, allWeights, "*"))

      values <- e$values[seq_len(M)]
      vectors <- e$vectors[,seq_len(M)]
    }
  }
  else
  {
    if(approx.eigen)
    {
      tmpSVD <- irlba::irlba(as.matrix(Matrix::tcrossprod(Z, Bchol)), nv = M)

      vectors <- Matrix::crossprod(Bchol, tmpSVD$v)
      values <- tmpSVD$d[seq_len(M)]^2
    }
    else
    {
      if(sum(npc) > 1000)
        warning("MFPCA with > 1000 univariate eigenfunctions and approx.eigen = FALSE. This may take some time...")

      e <- eigen(Matrix::crossprod(Bchol) %*% (stats::cov(allScores) * outer(allWeights, allWeights, "*")))

      values <- Re(e$values[seq_len(M)])
      vectors <- Re(e$vectors[,seq_len(M)])
    }
  }

  # normalization factors
  normFactors <- 1/sqrt(diag(as.matrix(Matrix::crossprod(Z %*% vectors))))

  ### Calculate scores
  scores <- Z %*% vectors * sqrt(N-1) # see defintion of Z above!
  scores <- as.matrix(scores %*% diag(sqrt(values) * normFactors, nrow = M, ncol = M)) # normalization

  ### Calculate eigenfunctions (incl. normalization)
  npcCum <- cumsum(c(0, npc)) # indices for blocks (-1)

  tmpWeights <- as.matrix(Matrix::crossprod(Z, Z %*%vectors))
  eFunctions <- foreach::foreach(j = seq_len(p)) %do% {
    univExpansion(type = type[j],
                  scores = 1/sqrt(weights[j] * values) * normFactors * t(tmpWeights[npcCum[j]+seq_len(npc[j]), , drop = FALSE]),
                  argvals = argvals[[j]],
                  functions = uniBasis[[j]]$functions,
                  params = uniBasis[[j]]$settings)
  }

  res <- list(values = values,
              functions = multiFunData(eFunctions),
              scores = scores,
              vectors = vectors,
              values = values,
              normFactors = normFactors,
              uniBasis = uniBasis,
              uniExpansions = uniExpansions
  )

  # calculate truncated Karhunen-Loeve representation (no mean here)
  if(fit)
    res$fit <- multivExpansion(multiFuns = res$functions, scores = scores)

  return(res)
}
```

### calcMFPCA_predict

```{r}
#' calcMFPCA_predict
#' @export
#' 
calcMFPCA_predict <- function(N, p, M, weights, npc, uniBasis,
                              normFactors_train, vectors_train, values_train)
{
  # combine all scores
  allScores <- foreach::foreach(j = seq_len(p), .combine = "cbind")%do%{uniBasis[[j]]$scores.pred}

  # block vector of weights
  allWeights <- foreach::foreach(j = seq_len(p), .combine = "c")%do%{rep(sqrt(weights[j]), npc[j])}

  Z <- allScores %*% Matrix::Diagonal(x = allWeights) / sqrt(N-1)

  ### Calculate scores
  scores <- Z %*% vectors_train * sqrt(N-1) # see defintion of Z above!
  scores <- as.matrix(scores %*% diag(sqrt(values_train) * normFactors_train, nrow = M, ncol = M)) # normalization

  return(scores)
}

```

### .PACE

```{r}
.PACE <- function(X, Y, Y.pred = NULL, nbasis = 10, pve = 0.99, npc = NULL, makePD = FALSE, cov.weight.type = "none")
{
  if (is.null(Y.pred))
    Y.pred = Y
  D = NCOL(Y)
  if(D != length(X)) # check if number of observation points in X & Y are identical
    stop("different number of (potential) observation points differs in X and Y!")
  I = NROW(Y)
  I.pred = NROW(Y.pred)
  d.vec = rep(X, each = I) # use given X-values for estimation of mu
  gam0 = mgcv::gam(as.vector(Y) ~ s(d.vec, k = nbasis))
  mu = mgcv::predict.gam(gam0, newdata = data.frame(d.vec = X))
  Y.tilde = Y - matrix(mu, I, D, byrow = TRUE)
  cov.sum = cov.count = cov.mean = matrix(0, D, D)
  for (i in seq_len(I)) {
    obs.points = which(!is.na(Y[i, ]))
    cov.count[obs.points, obs.points] = cov.count[obs.points, obs.points] + 1
    cov.sum[obs.points, obs.points] = cov.sum[obs.points, obs.points] + tcrossprod(Y.tilde[i, obs.points])
  }
  G.0 = ifelse(cov.count == 0, NA, cov.sum/cov.count)
  diag.G0 = diag(G.0)
  diag(G.0) = NA
  row.vec = rep(X, each = D) # use given X-values
  col.vec = rep(X, D) # use given X-values
  cov.weights <- switch(cov.weight.type,
                        none = rep(1, D^2),
                        counts = as.vector(cov.count),
                        stop("cov.weight.type ", cov.weight.type, " unknown in smooth covariance estimation"))

  npc.0 = matrix(mgcv::predict.gam(mgcv::gam(as.vector(G.0)~te(row.vec, col.vec, k = nbasis), weights = cov.weights),
                                   newdata = data.frame(row.vec = row.vec, col.vec = col.vec)), D, D)
  npc.0 = (npc.0 + t(npc.0))/2
  # no extra-option (useSymm) as in fpca.sc-method
  if (makePD) { # see fpca.sc
    npc.0 <- {
      tmp <- Matrix::nearPD(npc.0, corr = FALSE, keepDiag = FALSE,
                            do2eigen = TRUE, trace = options()$verbose)
      as.matrix(tmp$mat)
    }
  }

  ### numerical integration for calculation of eigenvalues (see Ramsay & Silverman, Chapter 8)
  w <- funData::.intWeights(X, method = "trapezoidal")
  Wsqrt <- diag(sqrt(w))
  Winvsqrt <- diag(1/(sqrt(w)))
  V <- Wsqrt %*% npc.0 %*% Wsqrt
  evalues = eigen(V, symmetric = TRUE, only.values = TRUE)$values
  ###
  evalues = replace(evalues, which(evalues <= 0), 0)
  npc = ifelse(is.null(npc), min(which(cumsum(evalues)/sum(evalues) > pve)), npc)
  efunctions = matrix(Winvsqrt%*%eigen(V, symmetric = TRUE)$vectors[, seq(len = npc)], nrow = D, ncol = npc)
  evalues = eigen(V, symmetric = TRUE, only.values = TRUE)$values[seq_len(npc)]  # use correct matrix for eigenvalue problem
  cov.hat = efunctions %*% tcrossprod(diag(evalues, nrow = npc, ncol = npc), efunctions)
  ### numerical integration for estimation of sigma2
  T.len <- X[D] - X[1] # total interval length
  T1.min <- min(which(X >= X[1] + 0.25*T.len)) # left bound of narrower interval T1
  T1.max <- max(which(X <= X[D] - 0.25*T.len)) # right bound of narrower interval T1
  DIAG = (diag.G0 - diag(cov.hat))[T1.min :T1.max] # function values
  # weights
  w <- funData::.intWeights(X[T1.min:T1.max], method = "trapezoidal")
  sigma2 <- max(1/(X[T1.max]-X[T1.min]) * sum(DIAG*w, na.rm = TRUE), 0) #max(1/T.len * sum(DIAG*w), 0)
  ####
  D.inv = diag(1/evalues, nrow = npc, ncol = npc)
  Z = efunctions
  Y.tilde = Y.pred - matrix(mu, I.pred, D, byrow = TRUE)
  fit = matrix(0, nrow = I.pred, ncol = D)
  scores = matrix(NA, nrow = I.pred, ncol = npc)
  # no calculation of confidence bands, no variance matrix
  for (i.subj in seq_len(I.pred)) {
    obs.points = which(!is.na(Y.pred[i.subj, ]))
    if (sigma2 == 0 & length(obs.points) < npc) {
      stop("Measurement error estimated to be zero and there are fewer observed points than PCs; scores cannot be estimated.")
    }
    Zcur = matrix(Z[obs.points, ], nrow = length(obs.points),
                  ncol = dim(Z)[2])
    ZtZ_sD.inv = solve(crossprod(Zcur) + sigma2 * D.inv)
    scores[i.subj, ] = ZtZ_sD.inv %*% crossprod(Zcur, Y.tilde[i.subj, obs.points])
    fit[i.subj, ] = t(as.matrix(mu)) + tcrossprod(scores[i.subj, ], efunctions)
  }
  ret.objects = c("fit", "scores", "mu", "efunctions", "evalues",
                  "npc", "sigma2") # add sigma2 to output
  ret = lapply(seq_len(length(ret.objects)), function(u) get(ret.objects[u]))
  names(ret) = ret.objects
  ret$estVar <- diag(cov.hat)
  return(ret)
}
```

### PACE

```{r}
PACE <- function(funDataObject, predData = NULL, nbasis = 10, pve = 0.99, npc = NULL, makePD = FALSE, cov.weight.type = "none")
{
  # check inputs
  if(! class(funDataObject) %in% c("funData", "irregFunData"))
    stop("Parameter 'funDataObject' must be a funData or irregFunData object.")
  if(dimSupp(funDataObject) != 1)
    stop("PACE: Implemented only for funData objects with one-dimensional support.")
  if(methods::is(funDataObject, "irregFunData")) # for irregular functional data, use funData representation
    funDataObject <- as.funData(funDataObject)

  if(is.null(predData))
    Y.pred = NULL # use only funDataObject
  else
  {
    if(!isTRUE(all.equal(funDataObject@argvals, predData@argvals)))
      stop("PACE: funDataObject and predData must be defined on the same domains!")

    Y.pred = predData@X
  }

  if(!all(is.numeric(nbasis), length(nbasis) == 1, nbasis > 0))
    stop("Parameter 'nbasis' must be passed as a number > 0.")

  if(!all(is.numeric(pve), length(pve) == 1, 0 <= pve, pve <= 1))
    stop("Parameter 'pve' must be passed as a number between 0 and 1.")

  if(!is.null(npc) & !all(is.numeric(npc), length(npc) == 1, npc > 0))
    stop("Parameter 'npc' must be either NULL or passed as a number > 0.")

  if(!is.logical(makePD))
    stop("Parameter 'makePD' must be passed as a logical.")

  if(!is.character(cov.weight.type))
    stop("Parameter 'cov.weight.type' must be passed as a character.")

  res <- .PACE(X = funDataObject@argvals[[1]],
               Y = funDataObject@X,
               Y.pred = Y.pred,
               nbasis = nbasis, pve = pve, npc = npc, makePD = makePD,
               cov.weight.type = cov.weight.type)

  return(list(mu = funData(funDataObject@argvals, matrix(res$mu, nrow = 1)),
              values = res$evalues,
              functions = funData(funDataObject@argvals, t(res$efunctions)),
              scores = res$scores,
              scores.pred = res$scores.pred,
              fit = funData(funDataObject@argvals, res$fit),
              npc = res$npc,
              sigma2 = res$sigma2,
              estVar = funData(funDataObject@argvals, matrix(res$estVar, nrow = 1))
  ))
}
```

### expandBasisFunction

```{r}
expandBasisFunction <- function(scores, argvals = functions@argvals, functions)
{
  if(dim(scores)[2] != nObs(functions))
    stop("expandBasisFunction: number of scores for each observation and number of eigenfunctions does not match.")

  # collapse higher-dimensional functions, multiply with scores and resize the result
  d <- dim(functions@X)
  nd <- length(d)

  if(nd == 2)
    resX <- scores %*% functions@X

  if(nd == 3)
  {
    resX <- array(NA, dim = c(dim(scores)[1], d[-1]))

    for(i in seq_len(d[2]))
      resX[,i,] <- scores %*% functions@X[,i,]
  }

  if(nd == 4)
  {
    resX <- array(NA, dim = c(dim(scores)[1], d[-1]))

    for(i in seq_len(d[2]))
      for(j in seq_len(d[3]))
          resX[,i,j,] <- scores %*% functions@X[,i,j,]
  }

  if(nd > 4) # slow solution due to aperm
  {
    resX <- aperm(plyr::aaply(.data = functions@X, .margins = 3:nd, 
                        .fun = function(x,y){y %*% x}, y = scores), 
                  c(nd-1,nd, seq_len((nd-2))))
    dimnames(resX) <- NULL
  }

  return( funData(argvals, resX) )
}
```

### univExpansion

```{r}
univExpansion <- function(type, scores, argvals = ifelse(!is.null(functions), functions@argvals, NULL), functions, params = NULL)
{
  # Parameter checking
  if(is.null(type))
    stop("Parameter 'type' is missing.")
  else
 {
   if(!is.character(type))
     stop("Parameter 'type' must be a character string. See ?univExpansion for details.")
 }

  if(is.null(scores))
    stop("Parameter 'scores' is missing.")
  else
    {
      if(!is.matrix(scores))
      stop("Parameter 'scores' must be passed as a matrix.")
    }

  if(is.numeric(argvals))
  {
    argvals <- list(argvals)
    warning("Parameter 'argvals' was passed as a vector and transformed to a list.")
  }

  if(is.null(functions))
  {
    if(is.null(argvals))
      stop("Must pass 'argvals' if 'functions' is NULL.")
    else
      {
        if(!is.list(argvals)) 
          stop("Parameter 'argvals' must be passed as a list.")
      }
  }
  else
  {
    if(class(functions) != "funData")
      stop("Parameter 'functions' must be a funData object.")

    # check interaction with other parameters
    if(nObs(functions) != NCOL(scores))
      stop("Number of scores per curve does not match the number of basis functions.")

    if(!is.null(argvals) & !isTRUE(all.equal(argvals, functions@argvals)))
      stop("The parameter 'argvals' does not match the argument values of 'functions'.")
  }

  if(!is.null(params) & !is.list(params))
    stop("The parameter 'params' must be passed as a list.")

  # start calculations
  params$scores <- scores
  params$functions <- functions

  if(is.numeric(argvals))
    argvals <- list(argvals)

  params$argvals <- argvals

  res <- switch(type,
                "given" = do.call(expandBasisFunction, params),
                "uFPCA" = do.call(expandBasisFunction, params),
                "UMPCA" = do.call(expandBasisFunction, params),
                "FCP_TPA" = do.call(expandBasisFunction, params),
                "splines1D" = do.call(splineFunction1D, params),
                "splines1Dpen" = do.call(splineFunction1D, params),
                "splines2D" = do.call(splineFunction2D, params),
                "splines2Dpen" = do.call(splineFunction2Dpen, params),
                "fda" = do.call(expandBasisFunction, params),
                "DCT2D" = do.call(dctFunction2D, params),
                "DCT3D" = do.call(dctFunction3D, params),
                "default" = do.call(expandBasisFunction, params),
                stop("Univariate Expansion for 'type' = ", type, " not defined!")
  )

  return(res)
}
```

## end-to-end MFPCA regression method

```{r}
# --- 1. MAIN FUNCTION: fit.mfpca_regression ---

#' End-to-End MFPCA Regression Pipeline
#'
#' Performs sample splitting, normalization, cross-validation (CV) to select
#' optimal functional (FPC) and scalar (SPC) components, linear regression fitting,
#' and test set prediction.
#'
#' @param W predictor_hybrid object containing all predictors.
#' @param y Vector. Response variable.
#' @param train_ratio Numeric scalar between 0 and 1 for splitting data (default 0.7).
#' @param n_fold Integer. Number of folds for cross-validation (default 5).
#' @param n_fpc_max Integer. Max number of FPCs to consider in CV (default 10).
#' @param n_spc_max Integer. Max number of SPCs to consider in CV (default 5).
#' @return List containing the final model fit, optimal parameters, and test MSE.
#' @export
fit.mfpca_regression <- function(W, y, train_ratio = 0.7, n_fold = 5, n_fpc_max = 10, n_spc_max = 5) {

  # --- 0. Setup and Initial Checks ---
  stopifnot(inherits(W, "predictor_hybrid"))

  # Standard parameters
  eval_point <- seq(0, 1, length.out = 100)

  uniExpansions <- list(
    list(type = "uFPCA", nbasis = 10, npc = n_fpc_max), 
    list(type = "uFPCA", nbasis = 10, npc = n_fpc_max)
  )

  # Adjust max components
  n_scalar <- W$n_scalar
  if (n_spc_max > n_scalar) n_spc_max <- n_scalar

  # --- 1. Split and Normalize Full Data ---

  split_obj <- split.hybrid(W, y, train_ratio)
  split_obj_normalized <- normailize.all(split_obj)

  # Extract components
  y_train_norm <- split_obj_normalized$response_train
  Z_train_norm <- split_obj_normalized$data_scalar_train
  f1_train_norm <- split_obj_normalized$data_fun_1_train
  f2_train_norm <- split_obj_normalized$data_fun_2_train
  N_train <- length(y_train_norm)

  # --- 2. Cross-Validation Setup ---

  cv_mat <- matrix(Inf, nrow = n_fpc_max, ncol = n_spc_max) 
  kfold_idx <- create_idx_kfold(N_train, n_fold)

  # --- 3. Cross-Validation Loop (Find Best n_fpc, n_spc) ---

  for (fold_num in 1:n_fold) {
    valid_idx <- kfold_idx[[fold_num]]
    train_idx_cv <- (1:N_train)[-valid_idx]

    # 3a. Prepare CV training and validation sets
    f1_train_cv <- f1_train_norm[train_idx_cv]
    f2_train_cv <- f2_train_norm[train_idx_cv]
    f1_valid_cv <- f1_train_norm[valid_idx]
    f2_valid_cv <- f2_train_norm[valid_idx]

    # Format functional data
    predictor_cv_train <- multiFunData(
      fd2funData(f1_train_cv, eval_point),
      fd2funData(f2_train_cv, eval_point)
    )
    predictor_cv_valid <- multiFunData(
      fd2funData(f1_valid_cv, eval_point),
      fd2funData(f2_valid_cv, eval_point)
    )

    Z_train_cv <- Z_train_norm[train_idx_cv, , drop = FALSE]
    Z_valid_cv <- Z_train_norm[valid_idx, , drop = FALSE]
    y_train_cv <- y_train_norm[train_idx_cv]
    y_valid_cv <- y_train_norm[valid_idx]

    # 3b. Fit PCA Models on CV Training Data
    MFPCA_fit_cv <- MFPCA(mFData = predictor_cv_train, mFData_predict = predictor_cv_valid, 
                          M = n_fpc_max, uniExpansions = uniExpansions)
    PCA_fit_cv <- prcomp(Z_train_cv, center = FALSE, scale = FALSE)

    FPC_scores_train_cv <- MFPCA_fit_cv$scores
    FPC_scores_valid_cv <- MFPCA_fit_cv$scores.pred
    PC_scores_train_cv <- PCA_fit_cv$x
    PC_scores_valid_cv <- predict(PCA_fit_cv, Z_valid_cv)

    # 3c. Inner Parameter Loop (Fit and Validate)
    for (n_fpc in 1:n_fpc_max) {
      for (n_spc in 1:n_spc_max) {

        # Select components
        FPC_train_select <- FPC_scores_train_cv[, 1:n_fpc, drop = FALSE]
        FPC_valid_select <- FPC_scores_valid_cv[, 1:n_fpc, drop = FALSE]
        PC_train_select <- PC_scores_train_cv[, 1:n_spc, drop = FALSE]
        PC_valid_select <- PC_scores_valid_cv[, 1:n_spc, drop = FALSE]

        # Combine selected scores for linear model
        train_data_lm <- data.frame(response = y_train_cv, FPC_train_select, PC_train_select)
        valid_data_lm <- data.frame(FPC_valid_select, PC_valid_select)

        # Fit Linear Regression Model
        linear_fit_cv <- lm(response ~ ., data = train_data_lm)

        # Predict, Calculate MSE, and Accumulate
        y_pred_valid <- predict(linear_fit_cv, valid_data_lm)
        cv_mat[n_fpc, n_spc] <- cv_mat[n_fpc, n_spc] + mean((y_pred_valid - y_valid_cv)^2)
      }
    }
  }

  # 4. Select Optimal Parameters
  min_idx <- arrayInd(which.min(cv_mat), dim(cv_mat))
  n_fpc_best <- min_idx[1]
  n_spc_best <- min_idx[2]

  # --- 5. Final Model Training and Prediction (Full Training Set) ---

  # Extract final components
  Z_test_norm <- split_obj_normalized$data_scalar_test
  f1_test_norm <- split_obj_normalized$data_fun_1_test
  f2_test_norm <- split_obj_normalized$data_fun_2_test
  y_test_norm <- split_obj_normalized$response_test

  # Prepare final functional data
  predictor_final_train <- multiFunData(fd2funData(f1_train_norm, eval_point), fd2funData(f2_train_norm, eval_point))
  predictor_final_test <- multiFunData(fd2funData(f1_test_norm, eval_point), fd2funData(f2_test_norm, eval_point))

  # Fit final PCA models
  MFPCA_fit_final <- MFPCA(mFData = predictor_final_train, mFData_predict = predictor_final_test, 
                          M = n_fpc_best, uniExpansions = uniExpansions)
  PCA_fit_final <- prcomp(Z_train_norm, center = FALSE, scale = FALSE)

  # Extract and combine final scores
  FPC_train_final <- MFPCA_fit_final$scores
  FPC_test_final <- MFPCA_fit_final$scores.pred
  PC_train_final <- PCA_fit_final$x[, 1:n_spc_best, drop = FALSE]
  PC_test_final <- predict(PCA_fit_final, Z_test_norm)[, 1:n_spc_best, drop = FALSE]

  train_data_final_lm <- data.frame(response = y_train_norm, FPC_train_final, PC_train_final)
  test_data_final_lm <- data.frame(FPC_test_final, PC_test_final)

  # Fit Final Linear Model
  final_linear_fit <- lm(response ~ ., data = train_data_final_lm)

  # Predict and Evaluate
  y_pred_test <- predict(final_linear_fit, test_data_final_lm)
  final_test_rmse <- sqrt(mean((y_pred_test - y_test_norm)^2))

  # --- 6. Return Results ---
  return(list(
    n_fpc_best = n_fpc_best,
    n_spc_best = n_spc_best,
    test_rmse = final_test_rmse,
    final_model = final_linear_fit,
    regression_coefficients = coef(final_linear_fit)
  ))
}
```

```{r}
load("C:/Users/jongmin/Documents/GitHub/fsPLS/simul/graphical_high_correlation/graphical_data.RData")
```

```{r}
library(refund)
eval_point<-seq(1, 100, length.out = 100)/100
n_iter_max = 40
n_basis_predictor = 10
n_scalar <- 9
library(foreach)
library(funData)
f1 = fd2funData(create_mock_fd(N, 1), eval_point)
f2 = fd2funData(create_mock_fd(N, 2), eval_point)
uniExpansions <- list(
  list(type = "uFPCA", nbasis = n_basis_predictor, npc = n_iter_max),
  list(type = "uFPCA", nbasis = n_basis_predictor, npc = n_iter_max)
)

n_fold = 5
n_scalar <- 9
rmse_simul_fpca <- matrix(NA, nrow = n_rep, ncol = 3)
colnames(rmse_simul_fpca) <- c("FPC+PC", "FPC_only", "PC_only")
```

```{r}
N <- 30
response <- c(1:N)
TRAIN_RATIO <- 0.7
create_mock_fd <- function(N_samples, seed) {
  set.seed(seed)
  mock_basis <- create.bspline.basis(rangeval = c(0, 1), nbasis = 20)
  C <- matrix(rnorm(20 * N_samples, mean = 10), nrow = 20)
  fd(C, mock_basis)
}
# 1. Setup Mock Hybrid Input
W_hybrid_mock <- predictor_hybrid(
  Z = matrix(1:(10*N), ncol = 10), # Simple sequential scalar data for verification
  functional_list = list(
    create_mock_fd(N, 1),
    create_mock_fd(N, 2)
  )
)

# 2. Split the data first
split_obj <- split_and_normailize.all(W_hybrid_mock, response, TRAIN_RATIO)

```
```

```{r}
    predictor_cv_train <- multiFunData(
f1, f2
    )
    MFPCA(mFData = predictor_cv_train, mFData_predict = predictor_cv_train, M = 5, uniExpansions = uniExpansions)
```

```{r}

  train_test_idx <- create_idx_train_test(n_sample, train.ratio)
  train_test_split <- split.all(data_fun_1[[rep_number]], data_fun_2[[rep_number]], data_scalar[[rep_number]], response[[rep_number]], train_test_idx)
  train_test_split_normalized <- split_and_normailize.all(train_test_split)

  kfold_idx <- create_idx_kfold(length(train_test_idx$idx_train), n_fold)
  cv_mat <- matrix(0, nrow = n_iter_max, ncol = n_scalar)
  for (fold_num in 1:n_fold){
    cat(paste0(fold_num, "th fold\n"))
    #train-valid split (5 fold)
    train_valid_idx <- kfold_idx[[fold_num]]
    cv_train_valid_split <- split.all(
      train_test_split$data_fun_1_train, train_test_split$data_fun_2_train, train_test_split$data_scalar_train, train_test_split$response_train,
      train_valid_idx)

    #normalizing
    cv_train_valid_split_normalized <- normailize.all(cv_train_valid_split)

    #format

    predictor_cv_valid <- multiFunData(
      fd2funData(cv_train_valid_split_normalized$data_fun_1_test, eval_point),
      fd2funData(cv_train_valid_split_normalized$data_fun_2_test, eval_point)
    )

    # train FPCA

    MFPCA_fit_cv <- MFPCA(mFData = predictor_cv_train, mFData_predict = predictor_cv_valid, M = n_iter_max, uniExpansions = uniExpansions)

    #train scalar PCA
    PCA_fit_train <- prcomp(cv_train_valid_split_normalized$data_scalar_train, center=F, scale = F)
    #validate
    for (n_fpc in 1:n_iter_max){
      for (n_spc in 1:n_scalar){
        cat(paste(n_fpc, n_spc, "\n"))
        FPC_score_cv <- (MFPCA_fit_cv$scores)[ ,(1:n_fpc)]
        PC_score_cv <- (PCA_fit_train$x)[ , (1:n_spc)]
        pc_score_cv_combined <- data.frame(cv_train_valid_split_normalized$response_train, FPC_score_cv, PC_score_cv)
        colnames(pc_score_cv_combined) <- c(
          c("response"),
          paste0( rep("FPC", n_fpc), as.character(1:n_fpc)),
          paste0( rep("PC", n_spc), as.character(1:n_spc))
        )

        # fit linear regression
        rank_now <- rankMatrix(pc_score_cv_combined[,-1])[1]
        print(rank_now)
        cat(paste("rank deficient:", (n_fpc + n_spc)-rank_now, "\n"))
        linear_fit_cv <- lm(response~., data = pc_score_cv_combined)

        #validation set pc scores
        PFC_score_valid <- (MFPCA_fit_cv$scores.pred)[ ,(1:n_fpc)]
        PC_score_valid <- (predict(PCA_fit_train, cv_train_valid_split_normalized$data_scalar_test))[ , (1:n_spc)]
        pc_score_valid_combined <- data.frame(PFC_score_valid, PC_score_valid)
        colnames(pc_score_valid_combined) <- c(
          paste0( rep("FPC", n_fpc), as.character(1:n_fpc)),
          paste0( rep("PC", n_spc), as.character(1:n_spc))
          )

        y_pred <- predict(linear_fit_cv, pc_score_valid_combined)
        cv_mat[n_fpc, n_spc] <- cv_mat[n_fpc, n_spc] +
          sqrt( mean( sum( (y_pred - cv_train_valid_split_normalized$response_test)^2 ) ))
        }
    }
  }
    n_iter_best <- arrayInd(which.min(cv_mat), dim(cv_mat))
    n_fpc_best  <- n_iter_best[1]
    n_spc_best  <- n_iter_best[2]
    print(n_fpc_best)
    print(n_spc_best)

    predictor_final_train <- multiFunData(
      fd2funData(train_test_split_normalized$data_fun_1_train, eval_point),
      fd2funData(train_test_split_normalized$data_fun_2_train, eval_point)
    )
    predictor_final_test <- multiFunData(
      fd2funData(train_test_split_normalized$data_fun_1_test, eval_point),
      fd2funData(train_test_split_normalized$data_fun_2_test, eval_point)
    )

    # train FPCA and scalar PCA
    MFPCA_fit_final <- MFPCA(mFData = predictor_final_train, mFData_predict = predictor_final_test, M = n_fpc_best, uniExpansions = uniExpansions)
    PCA_fit_final <- prcomp(train_test_split_normalized$data_scalar_train, center=F, scale = F)
    PFC_score_final_train <- MFPCA_fit_final$scores
    PC_score_final_train <- (predict(PCA_fit_final, train_test_split_normalized$data_scalar_train))[,c(1:n_spc_best)]
    pc_score_final_train_combined <- data.frame(train_test_split_normalized$response_train, PFC_score_final_train, PC_score_final_train)
    colnames(pc_score_final_train_combined) <- c(
      c("response"),
      paste0( rep("FPC", n_fpc_best), as.character(1:n_fpc_best)),
      paste0( rep("PC", n_spc_best), as.character(1:n_spc_best))
    )
    linear_fit_final <- lm(response~., data = pc_score_final_train_combined)

    # test set pc scores
    PFC_score_final_test <- MFPCA_fit_final$scores.pred
    PC_score_final_test <- (predict(PCA_fit_final, train_test_split_normalized$data_scalar_test))[,c(1:n_spc_best)]
    pc_score_final_test_combined <- data.frame(PFC_score_final_test, PC_score_final_test)
    colnames(pc_score_final_test_combined) <- c(
      paste0( rep("FPC", n_fpc_best), as.character(1:n_fpc_best)),
      paste0( rep("PC", n_spc_best), as.character(1:n_spc_best))
    )
    y_pred_final <- predict(linear_fit_final, pc_score_final_test_combined)
    rmse_simul_fpca[rep_number,1] <- sqrt(
      mean((y_pred_final - train_test_split_normalized$response_test)^2)
      )
    print(rmse_simul_fpca[rep_number,1])

    plot(-0.2:0.2, -0.2:0.2, pch = "", xlim = c(-0.2, 0.2),  ylim = c(-0.2, 0.2))
    plot(train_test_split_normalized$response_test, y_pred_final)
    abline(a=0, b=1)

    pc_score_final_train_combined_fpc_only <- data.frame(train_test_split_normalized$response_train, PFC_score_final_train)
    colnames(pc_score_final_train_combined_fpc_only) <- c(
      c("response"),
      paste0( rep("FPC", n_fpc_best), as.character(1:n_fpc_best))
    )
    linear_fit_final_fpc_only <- lm(response~., data = pc_score_final_train_combined_fpc_only)
    pc_score_final_test_combined_fpc_only <- data.frame(PFC_score_final_test)
    colnames(pc_score_final_test_combined_fpc_only) <- c(
      paste0( rep("FPC", n_fpc_best), as.character(1:n_fpc_best))
    )
    y_pred_final_fpc_only <- predict(linear_fit_final_fpc_only, pc_score_final_test_combined_fpc_only)
    rmse_simul_fpca[rep_number,2] <- sqrt(
      mean((y_pred_final_fpc_only - train_test_split_normalized$response_test)^2)
    )
    print(rmse_simul_fpca[rep_number,2])

    pc_score_final_train_combined_spc_only <- data.frame(train_test_split_normalized$response_train, PC_score_final_train)
    colnames(pc_score_final_train_combined_spc_only) <- c(
      c("response"),
      paste0( rep("PC", n_spc_best), as.character(1:n_spc_best))
    )
    linear_fit_final_spc_only <- lm(response~., data = pc_score_final_train_combined_spc_only)
    pc_score_final_test_combined_spc_only <- data.frame(PC_score_final_test)
    colnames(pc_score_final_test_combined_spc_only) <- c(
      paste0( rep("PC", n_spc_best), as.character(1:n_spc_best))
    )
    y_pred_final_spc_only <- predict(linear_fit_final_spc_only, pc_score_final_test_combined_spc_only)
    rmse_simul_fpca[rep_number,3] <- sqrt(
      mean((y_pred_final_spc_only - train_test_split_normalized$response_test)^2)
    )
    print(rmse_simul_fpca[rep_number,3])

```

## Documenting the package and building

We finish by running commands that will document, build, and install the package.  It may also be a good idea to check the package from within this file.

```{r}
litr::document() # <-- use instead of devtools::document()
# devtools::build()
# devtools::install()
# devtools::check(document = FALSE)
```