### 2025-12-30 simul.Rmd

---
title: "simulation"
output: html_document
date: "2025-12-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
# ==============================================================================
# EQUIVALENCE CHECK: exact_inprod() vs get_rho()
# ==============================================================================

library(fda)

# 1. DEFINE FUNCTIONS (To ensure environment is self-contained)
# ------------------------------------------------------------------------------

# A. Your Function (Logic: Matrix Multiplication)

# 2. PREPARE DATA
# ------------------------------------------------------------------------------
set.seed(123)
n_sample <- 50
n_basis  <- 10
basis <- create.bspline.basis(c(0,1), n_basis)

# Generate Random W (Data)
coefs_W <- matrix(rnorm(n_sample * n_basis), n_basis, n_sample)
Z_W <- matrix(rnorm(n_sample * 5), n_sample, 5)
fd_W <- fd(coefs_W, basis)
W_current <- list(
  Z = Z_W,
  functional_list = list(fd_W), # Single functional predictor for simplicity
  n_sample = n_sample,
  n_functional = 1,
  n_scalar = 5,
  n_basis_list = list(n_basis)
)

# CRITICAL STEP: Populate 'gram_list' for get_rho to work
# Your function assumes this exists. We must calculate it exactly as FDA does.
W_current$gram_list <- list(
  fda::inprod(basis, basis)
)

# Generate Random xi (Direction)
coefs_xi <- matrix(rnorm(n_basis), n_basis, 1) # 1 direction
Z_xi <- matrix(rnorm(5), 1, 5)
fd_xi <- fd(coefs_xi, basis)
xi_current <- list(
  Z = Z_xi,
  functional_list = list(fd_xi)
)

# 3. CONVERT INPUTS
# ------------------------------------------------------------------------------
# get_rho expects a flattened vector 'd_vec'.
# We must flatten xi_current: [functional_coefs | scalar_weights]
d_vec <- c(as.vector(coefs_xi), as.vector(Z_xi))

# 4. RUN COMPARISON
# ------------------------------------------------------------------------------

# Run My Function
val_exact <- exact_inprod(W_current, xi_current)

# Run Your Function (Note: d_vec is 1st arg, W is 2nd arg)
val_yours <- get_rho(d_vec, W_current)

# 5. VERDICT
# ------------------------------------------------------------------------------
diff <- max(abs(val_exact - val_yours))

cat("========================================\n")
cat("      EQUIVALENCE CHECK RESULT       \n")
cat("========================================\n")
cat(sprintf("Max Difference: %.10e\n", diff))

if(isTRUE(all.equal(val_exact, val_yours))) {
  cat("CONCLUSION: ✅ FUNCTIONS ARE IDENTICAL.\n")
  cat("Your implementation uses valid exact matrix algebra.\n")
} else {
  cat("CONCLUSION: ❌ FUNCTIONS DIFFER.\n")
}
```

```{r}
#' Exact Hybrid Inner Product (Matrix Based)
#' 
#' Calculates <W, xi> using Basis Inner Product Matrices (J).
#' Guarantees mathematical exactness for B-splines.
#' 
#' @param W Hybrid predictor (N samples)
#' @param xi Hybrid direction (1 sample/direction)
#' @return Numeric vector of length N (The scores)
exact_inprod <- function(W, xi) {

  # 1. Scalar Part (Dot Product)
  # (N x p) %*% (p x 1)
  scores <- as.matrix(W$Z) %*% t(as.matrix(xi$Z))

  # 2. Functional Part (Exact Integral via J matrix)
  for(k in seq_along(W$functional_list)) {
    fd_W  <- W$functional_list[[k]]
    fd_xi <- xi$functional_list[[k]]

    # Basis Inner Product Matrix: J[i,j] = Int(phi_i * phi_j)
    J <- fda::inprod(fd_W$basis, fd_W$basis)

    # Score = Coef_W^T * J * Coef_xi
    # (N x K) * (K x K) * (K x 1) -> (N x 1)

    # Efficient calculation:
    # We broadcast the (J %*% Coef_xi) vector across all N samples
    weighted_xi <- J %*% fd_xi$coefs

    # Add to scores (t(coefs) is N x K)
    scores <- scores + t(fd_W$coefs) %*% weighted_xi
  }

  return(as.vector(scores))
}
```

```{r}
# ==============================================================================
# SIMULATION: The Annihilation Property "Torture Test"
# Verifies: < W_resid[l+1], xi[l] >_H == 0 under extreme conditions
# ==============================================================================

library(fda)
library(ggplot2)
library(dplyr)
library(tidyr)

# ------------------------------------------------------------------------------
# 1. HELPER: EXACT INNER PRODUCT (Crucial for Verification)
# ------------------------------------------------------------------------------
exact_inprod <- function(W, xi) {
  # 1. Scalar Part
  scores <- as.matrix(W$Z) %*% t(as.matrix(xi$Z))

  # 2. Functional Part (Exact Integral via J matrix)
  for(k in seq_along(W$functional_list)) {
    fd_W  <- W$functional_list[[k]]
    fd_xi <- xi$functional_list[[k]]
    J <- fda::inprod(fd_W$basis, fd_W$basis)
    weighted_xi <- J %*% fd_xi$coefs
    scores <- scores + t(fd_W$coefs) %*% weighted_xi
  }
  return(as.vector(scores))
}

# ------------------------------------------------------------------------------
# 2. DATA GENERATION: "The Torture Chamber"
# ------------------------------------------------------------------------------
generate_torture_data <- function(n_sample = 100) {
  # A. Shared Latent Factors (Creates collinearity)
  t1 <- rnorm(n_sample, sd=10)
  t2 <- rnorm(n_sample, sd=0.1) # Tiny variance component

  eval_points <- seq(0, 1, length.out=100)
  basis <- create.bspline.basis(c(0,1), 15) # Higher basis dim

  # B. Functional Predictor 1: High Magnitude, Low Freq
  X1 <- outer(t1, sin(2*pi*eval_points)) 
  fd1 <- Data2fd(eval_points, t(X1), basis)

  # C. Functional Predictor 2: Low Magnitude, High Freq (Noise trap)
  X2 <- outer(t2, sin(10*pi*eval_points)) + matrix(rnorm(n_sample*100, sd=0.01), 100, n_sample)
  fd2 <- Data2fd(eval_points, t(X2), basis)

  # D. Scalars: 50 Variables, Rank Deficient (Rank ~ 2)
  Z_base <- matrix(rnorm(n_sample * 2), n_sample, 2)
  Z_noise <- matrix(rnorm(n_sample * 50, sd=0.001), n_sample, 50)
  # Mix: All 50 scalars depend on just 2 random vectors
  Z <- Z_base %*% matrix(rnorm(2*50), 2, 50) + Z_noise

  # E. Response
  y <- t1 * 0.5 + t2 * 10 + rnorm(n_sample)

  W_init <- predictor_hybrid(Z, list(fd1, fd2), eval_points)
  return(list(W = W_init, y = y))
}

# ------------------------------------------------------------------------------
# 3. RUN SIMULATION
# ------------------------------------------------------------------------------
set.seed(999)
data <- generate_torture_data(100)
processed <- split_and_normalize.all(data$W, data$y, train_ratio=0.99)

W_current <- processed$predictor_train
y_current <- processed$response_train

# TEST PARAMETERS
# We use different lambdas for different functions to stress the metric
lambda_vec <- c(10, 0.01) 
n_iter <- 5
results_df <- data.frame()

message("Running Torture Test (Exact Matrix Mode)...")

for(l in 1:n_iter) {

  # 1. Get Direction (Xi) - Penalized
  xi_current <- get_xi_hat_linear_pen(W_current, y_current, lambda_vec)

  # 2. Get Scores (Rho) -- EXACT
  #rho_current <- exact_inprod(W_current, xi_current)
  rho_current <- get_rho(W_current, xi_current)

  # 3. Get Loadings (Delta)
  delta_current <- get_delta(W_current, rho_current) 
  nu_current    <- get_nu(y_current, rho_current)

  # 4. MEASURE SIGNAL BEFORE DEFLATION
  signal_mag <- sqrt(sum(exact_inprod(W_current, xi_current)^2))

  # 5. Deflate
  W_next <- residualize_predictor(W_current, rho_current, delta_current)
  y_next <- residualize_y(y_current, rho_current, nu_current)

  # 6. MEASURE LEAKAGE AFTER DEFLATION (The Proof Condition)
  leakage_vec <- exact_inprod(W_next, xi_current)
  leakage_mag <- sqrt(sum(leakage_vec^2))

  # 7. Log
  status <- ifelse(leakage_mag < 1e-9, "PASS", "FAIL")

  results_df <- rbind(results_df, data.frame(
    Iteration = l,
    Signal_Strength = signal_mag,
    Residual_Leakage = leakage_mag,
    Status = status
  ))

  cat(sprintf("Iter %d: Signal=%.2e | Leakage=%.2e | %s\n", 
              l, signal_mag, leakage_mag, status))

  W_current <- W_next
  y_current <- y_next
}

# ------------------------------------------------------------------------------
# 4. VISUALIZATION
# ------------------------------------------------------------------------------
cat("\n=======================================================\n")
cat("      TORTURE TEST RESULTS \n")
cat("=======================================================\n")
print(results_df)

df_long <- tidyr::pivot_longer(results_df, cols = c("Signal_Strength", "Residual_Leakage"))

ggplot(df_long, aes(x=factor(Iteration), y=value, fill=name)) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_y_continuous(trans="log10") +
  labs(title="Annihilation Torture Test",
       subtitle="Even with high collinearity and penalties, leakage must be ~0",
       y="Magnitude (Log10)",
       x="Iteration") +
  theme_minimal() +
  scale_fill_manual(values=c("firebrick", "steelblue")) +
  geom_hline(yintercept=1e-10, linetype="dashed")
```

```{r}
# ==============================================================================
# STUDY: PLS Geometric Properties (Orthogonality & Relevance)
# ==============================================================================

library(fda)
library(ggplot2)
library(dplyr)
library(tidyr)

# ------------------------------------------------------------------------------
# 1. SETUP
# ------------------------------------------------------------------------------

lambda_scenarios <- list(
  "Scenario A: weak"    = c(0.1, 0.1),         # Pure NIPALS
  "Scenario B: mixed"  = c(0.1, 5),    # Mixed Smoothness
  "Scenario C: strong"   = c(5, 5)        # Strong Regularization
)

n_rep <- 100       # Repetitions
n_comp <- 6        # Components
results_ortho <- data.frame() # Store orthogonality results
results_corr  <- data.frame() # Store correlation results

# Data Generation Function
generate_robust_data <- function(n_sample = 200, n_scalar = 50) {
  t1 <- rnorm(n_sample, sd = 10)
  t2 <- rnorm(n_sample, sd = 5)
  t3 <- rnorm(n_sample, sd = 2)

  eval_points <- seq(0, 1, length.out = 100)
  basis <- create.bspline.basis(c(0, 1), 10)

  X1 <- outer(t1, sin(2*pi*eval_points)) + outer(t2, cos(2*pi*eval_points))
  fd1 <- Data2fd(eval_points, t(X1), basis)

  X2 <- outer(t1, sin(4*pi*eval_points)) + outer(t2, cos(4*pi*eval_points))
  fd2 <- Data2fd(eval_points, t(X2), basis)

  Z <- matrix(0, n_sample, n_scalar)
  for(j in 1:n_scalar) Z[,j] <- t1*runif(1) + t3*runif(1) + rnorm(n_sample)

  y <- t1 * 2 - t2 * 1.5 + rnorm(n_sample)
  W <- predictor_hybrid(Z, list(fd1, fd2), eval_points)

  return(list(W = W, y = y))
}

# Helper: Compute Inner Product of Two Hybrid Vectors
# (Required to check orthogonality of xi)
hybrid_inner_product <- function(h1, h2) {
  # Scalar dot product
  val <- sum(h1$Z * h2$Z)
  # Functional inner products
  for(k in seq_along(h1$functional_list)) {
    val <- val + inprod(h1$functional_list[[k]], h2$functional_list[[k]])
  }
  return(as.numeric(val))
}

# ------------------------------------------------------------------------------
# 2. RUN SIMULATION LOOP
# ------------------------------------------------------------------------------
message("Starting Extended Simulation...")
set.seed(999)

for (scenario_name in names(lambda_scenarios)) {
  lam_pair <- lambda_scenarios[[scenario_name]]

  for (i in 1:n_rep) {
    # Generate & Fit
    data <- generate_robust_data(n_sample = 200)
    processed <- split_and_normalize.all(data$W, data$y, train_ratio = 0.8)

    fit <- fit.hybridPLS(
      W = processed$predictor_train, 
      y = processed$response_train, 
      n_iter = n_comp, 
      lambda = lam_pair
    )

    # --- METRIC 1: SCORE (Rho) ORTHOGONALITY ---
    rho_mat <- do.call(cbind, fit$rho)
    cor_rho <- cor(rho_mat)
    diag(cor_rho) <- 0
    max_rho_overlap <- max(abs(cor_rho))

    # --- METRIC 2: DIRECTION (Xi) ORTHOGONALITY ---
    # We calculate the max pairwise inner product between different weight vectors
    # Note: xi are normalized, so self-inner-product is 1.
    xi_list <- fit$xi
    xi_overlaps <- numeric()

    for(j in 1:(n_comp-1)) {
      for(k in (j+1):n_comp) {
        # Dot product between direction j and direction k
        dp <- hybrid_inner_product(xi_list[[j]], xi_list[[k]])
        xi_overlaps <- c(xi_overlaps, abs(dp))
      }
    }
    max_xi_overlap <- max(xi_overlaps)

    # Record Orthogonality Stats
    results_ortho <- rbind(results_ortho, data.frame(
      Scenario = scenario_name,
      Repetition = i,
      Max_Score_Corr = max_rho_overlap,    # Should be ~0
      Max_Direction_Cos = max_xi_overlap   # Can be > 0
    ))

    # --- METRIC 3: SCORE vs RESPONSE CORRELATION ---
    # Does Comp 1 explain Y better than Comp 2?
    y_vec <- processed$response_train
    cor_y <- cor(rho_mat, y_vec)

    # Store trajectory for this repetition
    for(k in 1:n_comp) {
      results_corr <- rbind(results_corr, data.frame(
        Scenario = scenario_name,
        Repetition = i,
        Component = k,
        Correlation_with_Y = abs(cor_y[k, 1])
      ))
    }
  }
}

# ------------------------------------------------------------------------------
# 3. REPORT GENERATION
# ------------------------------------------------------------------------------

# --- TABLE 1: Orthogonality (Scores vs Directions) ---
summary_ortho <- results_ortho %>%
  group_by(Scenario) %>%
  summarise(
    Score_Orthogonality = mean(Max_Score_Corr),
    Direction_Overlap   = mean(Max_Direction_Cos)
  )

cat("\n=================================================================\n")
cat(" TABLE 1: Geometric Properties (Scores vs Directions)\n")
cat("=================================================================\n")
cat("NOTE: 'Score' should be 0 (Orthogonal). 'Direction' can be > 0.\n\n")
print(summary_ortho)

# --- VISUALIZATION: Relevance Decay ---
# We want to see if correlation drops as component index increases

summary_corr <- results_corr %>%
  group_by(Scenario, Component) %>%
  summarise(
    Mean_Corr = mean(Correlation_with_Y),
    SD_Corr = sd(Correlation_with_Y)
  )

p_decay <- ggplot(summary_corr, aes(x = Component, y = Mean_Corr, color = Scenario)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = Mean_Corr - SD_Corr, ymax = Mean_Corr + SD_Corr), width = 0.2) +
  labs(title = "Predictive Relevance Decay",
       subtitle = "Correlation between PLS Scores and Response Y",
       y = "|Correlation(Score, Y)|",
       x = "Component Index") +
  theme_bw() +
  scale_x_continuous(breaks = 1:n_comp)

print(p_decay)
```

```{r}
# ==============================================================================
# PLS VERIFICATION: Prediction & Beta Recovery
# ==============================================================================

library(fda)
library(ggplot2)
library(gridExtra)
library(FSHybridPLS) # Your package

# ------------------------------------------------------------------------------
# 1. GENERATE TRUTH (Linear Model)
# ------------------------------------------------------------------------------
generate_regression_truth <- function(n_sample = 200, n_eval = 100, noise_sd = 0.5) {

  eval_points <- seq(0, 1, length.out = n_eval)
  basis_obj <- create.bspline.basis(c(0, 1), 20)

  # --- A. True Coefficient Function (Beta) ---
  # A distinct shape: Sine wave with a trend
  true_beta_vec <- sin(2 * pi * eval_points) + 1.5 * eval_points
  # Center it (standard practice)
  true_beta_vec <- true_beta_vec - mean(true_beta_vec)

  # --- B. Functional Predictors (X) ---
  # Smooth random curves
  coefs <- matrix(rnorm(20 * n_sample), 20, n_sample)
  # Smooth the coefficients to make curves realistic
  smooth_mat <- chol(ar1_cov(20, 0.8))
  coefs <- t(smooth_mat) %*% coefs

  fd_obj <- fd(coefs, basis_obj)
  X_vals <- eval.fd(eval_points, fd_obj) # 100 x 200

  # --- C. Response (Y) ---
  # Y = Integral( X(t) * Beta(t) ) + Noise
  # Approx integral as dot product * dt
  dt <- 1 / n_eval

  # Calculate signal
  # (100 x 200) * (100 x 1) -> broadcasting requires care or loop
  # Signal = colSums( X * beta ) * dt
  signal <- colSums(X_vals * true_beta_vec) * dt

  # Add noise
  y <- signal + rnorm(n_sample, sd = noise_sd)

  # Scalar Z (Noise only, to prove PLS ignores it)
  Z <- matrix(rnorm(n_sample), n_sample, 1)

  W <- predictor_hybrid(Z, list(fd_obj), eval_points)

  return(list(
    W = W, y = y, 
    true_beta = true_beta_vec,
    eval_points = eval_points
  ))
}

# ------------------------------------------------------------------------------
# 2. RUN SIMULATION
# ------------------------------------------------------------------------------
set.seed(42)
# Generate
sim_data <- generate_regression_truth(n_sample = 200, noise_sd = 0.1)

# Split (Train/Test)
processed <- split_and_normalize.all(sim_data$W, sim_data$y, train_ratio = 0.7)

# Fit PLS
# Using 4 components (should be enough to capture the beta shape)
# Lambda = 0 ensures we test the core algorithm, not the penalty
fit <- fit.hybridPLS(
  W = processed$predictor_train, 
  y = processed$response_train, 
  n_iter = 4, 
  lambda = c(0),
  validation_data = list(W_test = processed$predictor_test, y_test = processed$response_test)
)

# ------------------------------------------------------------------------------
# 3. VERIFICATION PLOTS
# ------------------------------------------------------------------------------

# --- A. Prediction Check (Test Set) ---
# Reconstruct Y_pred for Test Set using the Final Beta (Comp 4)
beta_final <- fit$beta[[4]]
y_pred <- inprod.predictor_hybrid(processed$predictor_test, beta_final)
y_true <- processed$response_test

r2_test <- 1 - sum((y_true - y_pred)^2) / sum((y_true - mean(y_true))^2)

df_pred <- data.frame(Observed = y_true, Predicted = y_pred)

p1 <- ggplot(df_pred, aes(x = Observed, y = Predicted)) +
  geom_point(alpha = 0.6, color = "blue") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  labs(title = paste0("Prediction Accuracy (R2 = ", round(r2_test, 3), ")"),
       subtitle = "Points should lie on the red diagonal") +
  theme_bw()

# --- B. Beta Recovery Check ---
# Extract estimated functional beta
beta_est_fd <- fit$beta[[4]]$functional_list[[1]]
beta_est_vec <- eval.fd(sim_data$eval_points, beta_est_fd)

# Normalize both to compare SHAPE (scale might differ due to normalization)
scale_vec <- function(x) as.vector((x - mean(x)) / sd(x))

df_beta <- data.frame(
  Time = sim_data$eval_points,
  True = scale_vec(sim_data$true_beta),
  Estimated = scale_vec(beta_est_vec)
)

beta_corr <- cor(df_beta$True, df_beta$Estimated)

p2 <- ggplot(df_beta, aes(x = Time)) +
  geom_line(aes(y = True, color = "True Beta"), linewidth = 1.2, linetype = "dashed") +
  geom_line(aes(y = Estimated, color = "PLS Estimate"), linewidth = 1) +
  scale_color_manual(values = c("True Beta" = "black", "PLS Estimate" = "red")) +
  labs(title = paste0("Beta Reconstruction (Corr = ", round(beta_corr, 3), ")"),
       subtitle = "Red line should match Black dashed line") +
  theme_bw() + theme(legend.position = "bottom")

# ------------------------------------------------------------------------------
# 4. RESULTS
# ------------------------------------------------------------------------------
grid.arrange(p1, p2, nrow = 1)

cat("\n========================================\n")
cat("      PLS REGRESSION VERIFICATION       \n")
cat("========================================\n")
cat(sprintf("1. Test R-Squared:      %.4f  [%s]\n", r2_test, ifelse(r2_test > 0.9, "PASS", "FAIL")))
cat(sprintf("2. Beta Shape Corr:     %.4f  [%s]\n", beta_corr, ifelse(beta_corr > 0.95, "PASS", "FAIL")))

if(r2_test > 0.9 && beta_corr > 0.95) {
  cat("\nCONCLUSION: ✅ PLS Works. It predicts accurately and learns the correct shape.\n")
} else {
  cat("\nCONCLUSION: ❌ Something is wrong.\n")
}
```

```{r}
# ==============================================================================
# PLS VERIFICATION: Prediction & Beta Recovery
# ==============================================================================

library(fda)
library(ggplot2)
library(gridExtra)
library(FSHybridPLS) # Your package

# ------------------------------------------------------------------------------
# 1. GENERATE TRUTH (Linear Model)
# ------------------------------------------------------------------------------
generate_regression_truth <- function(n_sample = 200, n_eval = 100, noise_sd = 0.01) {

  eval_points <- seq(0, 1, length.out = n_eval)
  basis_obj <- create.bspline.basis(c(0, 1), 20)

  # --- A. True Coefficient Function (Beta) ---
  # A distinct shape: Sine wave with a trend
  true_beta_vec <- sin(2 * pi * eval_points) + 1.5 * eval_points
  # Center it (standard practice)
  true_beta_vec <- true_beta_vec - mean(true_beta_vec)

  # --- B. Functional Predictors (X) ---
  # Smooth random curves
  coefs <- matrix(rnorm(20 * n_sample), 20, n_sample)
  # Smooth the coefficients to make curves realistic
  smooth_mat <- chol(ar1_cov(20, 0.8))
  coefs <- t(smooth_mat) %*% coefs

  fd_obj <- fd(coefs, basis_obj)
  X_vals <- eval.fd(eval_points, fd_obj) # 100 x 200

  # --- C. Response (Y) ---
  # Y = Integral( X(t) * Beta(t) ) + Noise
  # Approx integral as dot product * dt
  dt <- 1 / n_eval

  # Calculate signal
  # (100 x 200) * (100 x 1) -> broadcasting requires care or loop
  # Signal = colSums( X * beta ) * dt
  signal <- colSums(X_vals * true_beta_vec) * dt

  # Add noise
  y <- signal + rnorm(n_sample, sd = noise_sd)

  # Scalar Z (Noise only, to prove PLS ignores it)
  Z <- matrix(rnorm(n_sample), n_sample, 1)

  W <- predictor_hybrid(Z, list(fd_obj), eval_points)

  return(list(
    W = W, y = y, 
    true_beta = true_beta_vec,
    eval_points = eval_points
  ))
}

# ------------------------------------------------------------------------------
# 2. RUN SIMULATION
# ------------------------------------------------------------------------------
set.seed(42)
# Generate
sim_data <- generate_regression_truth(n_sample = 200, noise_sd = 0.01)

# Split (Train/Test)
processed <- split_and_normalize.all(sim_data$W, sim_data$y, train_ratio = 0.7)

# Fit PLS
# Using 4 components (should be enough to capture the beta shape)
# Lambda = 0 ensures we test the core algorithm, not the penalty
fit <- fit.hybridPLS(
  W = processed$predictor_train, 
  y = processed$response_train, 
  n_iter = 4, 
  lambda = c(0),
  validation_data = list(W_test = processed$predictor_test, y_test = processed$response_test)
)

# ------------------------------------------------------------------------------
# 3. VERIFICATION PLOTS
# ------------------------------------------------------------------------------

# --- A. Prediction Check (Test Set) ---
# Reconstruct Y_pred for Test Set using the Final Beta (Comp 4)
beta_final <- fit$beta[[4]]
y_pred <- inprod.predictor_hybrid(processed$predictor_test, beta_final)
y_true <- processed$response_test

r2_test <- 1 - sum((y_true - y_pred)^2) / sum((y_true - mean(y_true))^2)

df_pred <- data.frame(Observed = y_true, Predicted = y_pred)

p1 <- ggplot(df_pred, aes(x = Observed, y = Predicted)) +
  geom_point(alpha = 0.6, color = "blue") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  labs(title = paste0("Prediction Accuracy (R2 = ", round(r2_test, 3), ")"),
       subtitle = "Points should lie on the red diagonal") +
  theme_bw()

# --- B. Beta Recovery Check ---
# Extract estimated functional beta
beta_est_fd <- fit$beta[[4]]$functional_list[[1]]
beta_est_vec <- eval.fd(sim_data$eval_points, beta_est_fd)

# Normalize both to compare SHAPE (scale might differ due to normalization)
scale_vec <- function(x) as.vector((x - mean(x)) / sd(x))

df_beta <- data.frame(
  Time = sim_data$eval_points,
  True = scale_vec(sim_data$true_beta),
  Estimated = scale_vec(beta_est_vec)
)

beta_corr <- cor(df_beta$True, df_beta$Estimated)

p2 <- ggplot(df_beta, aes(x = Time)) +
  geom_line(aes(y = True, color = "True Beta"), linewidth = 1.2, linetype = "dashed") +
  geom_line(aes(y = Estimated, color = "PLS Estimate"), linewidth = 1) +
  scale_color_manual(values = c("True Beta" = "black", "PLS Estimate" = "red")) +
  labs(title = paste0("Beta Reconstruction (Corr = ", round(beta_corr, 3), ")"),
       subtitle = "Red line should match Black dashed line") +
  theme_bw() + theme(legend.position = "bottom")

# ------------------------------------------------------------------------------
# 4. RESULTS
# ------------------------------------------------------------------------------
grid.arrange(p1, p2, nrow = 1)

cat("\n========================================\n")
cat("      PLS REGRESSION VERIFICATION       \n")
cat("========================================\n")
cat(sprintf("1. Test R-Squared:      %.4f  [%s]\n", r2_test, ifelse(r2_test > 0.9, "PASS", "FAIL")))
cat(sprintf("2. Beta Shape Corr:     %.4f  [%s]\n", beta_corr, ifelse(beta_corr > 0.95, "PASS", "FAIL")))

if(r2_test > 0.9 && beta_corr > 0.95) {
  cat("\nCONCLUSION: ✅ PLS Works. It predicts accurately and learns the correct shape.\n")
} else {
  cat("\nCONCLUSION: ❌ Something is wrong.\n")
}
```

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

## is normalization hurtful?
```{r}
# ==============================================================================
# DIAGNOSTIC: Is Normalization Killing PLS?
# ==============================================================================

library(fda)
library(Matrix)
library(ggplot2)

# 1. GENERATE LATENT DATA (Signal defined by Variance)
# ------------------------------------------------------------------------------
# We create 2 components:
# Comp 1: Huge Variance (SD=10) -> Should be found FIRST
# Comp 2: Small Variance (SD=2) -> Should be found SECOND
generate_variance_signal <- function(n_sample = 200) {
  t1 <- rnorm(n_sample, sd = 10)
  t2 <- rnorm(n_sample, sd = 2)

  # Orthogonalize exactly
  T_mat <- svd(cbind(t1, t2))$u %*% diag(c(10, 2))

  # Functional Predictor X (Evaluation points 0 to 1)
  eval_points <- seq(0, 1, length.out=100)
  shape1 <- sin(2*pi*eval_points)
  shape2 <- cos(2*pi*eval_points)

  X <- outer(T_mat[,1], shape1) + outer(T_mat[,2], shape2) + rnorm(n_sample*100, sd=0.1)

  # Scalar Predictor Z (Just noise for now, to isolate Functional issue)
  Z <- matrix(rnorm(n_sample * 5), n_sample, 5)

  # Response Y (Driven by T1 mostly)
  y <- T_mat[,1] * 1 + T_mat[,2] * 1 + rnorm(n_sample, sd=0.1)

  # Create Objects
  basis_obj <- create.bspline.basis(c(0, 1), 20)
  fd_obj <- Data2fd(eval_points, t(X), basis_obj)

  W <- predictor_hybrid(Z, list(fd_obj), eval_points)

  return(list(W = W, y = y, true_scores = T_mat))
}

set.seed(123)
data <- generate_variance_signal(200)

# ==============================================================================
# SCENARIO A: RAW DATA (Preserve Variance)
# ==============================================================================
message("\n--- Running Scenario A: RAW DATA (Variance Preserved) ---")

# Manual Centering (Required for PLS, but NO scaling)
y_centered <- data$y - mean(data$y)
W_centered <- data$W
# (Assuming your PLS centers internally or we can pass centered data. 
#  For strictness, we assume Z is roughly centered by generation.)

fit_raw <- fit.hybridPLS(
  W = W_centered, 
  y = y_centered, 
  n_iter = 2, 
  lambda = c(0) # No penalty to isolate normalization issue
)

rho_raw <- do.call(cbind, fit_raw$rho)
cor_raw <- cor(rho_raw, data$true_scores)
print("Correlation (Estimated vs True):")
print(round(cor_raw, 3))

# ==============================================================================
# SCENARIO B: NORMALIZED DATA (Variance Destroyed)
# ==============================================================================
message("\n--- Running Scenario B: NORMALIZED DATA (Variance = 1) ---")

# Run your normalization pipeline
# We use train_ratio = 0.99 just to keep most data for the check
processed <- split_and_normalize.all(data$W, data$y, train_ratio = 0.99)

# Extract the TRUE scores corresponding to the training set indices
train_idx <- processed$details$split_indices$train
true_scores_subset <- data$true_scores[train_idx, ]

fit_norm <- fit.hybridPLS(
  W = processed$predictor_train, 
  y = processed$response_train, 
  n_iter = 2, 
  lambda = c(0)
)

rho_norm <- do.call(cbind, fit_norm$rho)
cor_norm <- cor(rho_norm, true_scores_subset)
print("Correlation (Estimated vs True):")
print(round(cor_norm, 3))

# ==============================================================================
# RESULT INTERPRETATION
# ==============================================================================
# Look at the diagonal of the correlation matrices.
# Ideally: [1,1] > 0.9 (Comp 1 found), [2,2] > 0.9 (Comp 2 found).

score_1_raw <- abs(cor_raw[1,1])
score_1_norm <- abs(cor_norm[1,1])

cat("\n========================================\n")
cat("          DIAGNOSTIC RESULTS            \n")
cat("========================================\n")
cat(paste0("Scenario A (Raw) Score 1 Recovery: ", round(score_1_raw, 3), "\n"))
cat(paste0("Scenario B (Norm) Score 1 Recovery: ", round(score_1_norm, 3), "\n"))

if (score_1_raw > 0.9 && score_1_norm < 0.9) {
  cat("\nCONCLUSION: 🚨 Normalization is the problem.\n")
  cat("By forcing unit variance, 'split_and_normalize.all' destroys the \n")
  cat("variance structure that PLS uses to identify the primary component.\n")
} else if (score_1_raw < 0.9) {
  cat("\nCONCLUSION: ⚠️  The Algorithm is the problem.\n")
  cat("Even with raw data, PLS failed to find the latent component.\n")
} else {
  cat("\nCONCLUSION: ✅ Both methods worked? (Check if simulation noise was too low)\n")
}
```

##

```{r}

# ==============================================================================
# HYBRID PLS VERIFICATION: Latent Factor & Coefficient Recovery
# ==============================================================================

library(fda)
library(ggplot2)
library(gridExtra) # For arranging plots
library(FSHybridPLS) # Your package

# ------------------------------------------------------------------------------
# 1. DATA GENERATION: The "Ground Truth"
# ------------------------------------------------------------------------------
generate_verification_data <- function(n_sample = 200, n_eval = 100) {

  eval_points <- seq(0, 1, length.out = n_eval)

  # --- A. Latent Scores (Truth) ---
  # We create 2 orthogonal components with distinct variances
  # Comp 1: High Variance (SD=10) -> The primary signal
  # Comp 2: Lower Variance (SD=4) -> The secondary signal
  set.seed(123)
  t1 <- rnorm(n_sample, sd = 10)
  t2 <- rnorm(n_sample, sd = 4)
  # Force exact orthogonality for clean testing
  T_mat <- svd(cbind(t1, t2))$u %*% diag(c(10, 4))

  # --- B. Loadings (Shapes) ---
  # Shape 1: Low frequency (Sine wave)
  # Shape 2: High frequency (Cosine wave)
  shape1 <- sin(2 * pi * eval_points)
  shape2 <- cos(4 * pi * eval_points)

  # --- C. Functional Predictor X ---
  # X = T * P' + Noise
  X_clean <- outer(T_mat[,1], shape1) + outer(T_mat[,2], shape2)
  X_noise <- matrix(rnorm(n_sample * n_eval, sd = 0.5), n_sample, n_eval)
  X <- X_clean + X_noise

  # --- D. Response Y ---
  # Y depends on both components: Y = 1*t1 - 2*t2 + Noise
  # This implies the TRUE BETA shape is roughly: (1*shape1) - (2*shape2)
  y <- T_mat[,1] * 1 + T_mat[,2] * (-2) + rnorm(n_sample, sd = 1)

  # Calculate True Beta Shape (Theoretical)
  # (Note: Normalized to SD=1 for shape comparison)
  true_beta_raw <- (1 * shape1) + (-2 * shape2)
  true_beta_norm <- (true_beta_raw - mean(true_beta_raw))/sd(true_beta_raw)

  # --- E. Objects ---
  basis_obj <- create.bspline.basis(c(0, 1), 20)
  fd_obj <- Data2fd(eval_points, t(X), basis_obj)

  # Dummy scalar (not used for signal, just to satisfy object structure)
  Z <- matrix(rnorm(n_sample), n_sample, 1)

  W <- predictor_hybrid(Z, list(fd_obj), eval_points)

  return(list(
    W = W, y = y, 
    true_scores = T_mat, 
    true_beta = true_beta_norm,
    eval_points = eval_points
  ))
}

# ------------------------------------------------------------------------------
# 2. EXECUTION
# ------------------------------------------------------------------------------
message("Generating Data...")
sim_data <- generate_verification_data(n_sample = 200)

message("Processing Data (Split & Normalize)...")
# We use a large training ratio to ensure stable estimation for verification
processed <- split_and_normalize.all(sim_data$W, sim_data$y, train_ratio = 0.8)

# Extract True Scores corresponding to the Training Set
train_idx <- processed$details$split_indices$train
true_scores_train <- sim_data$true_scores[train_idx, ]

message("Fitting Hybrid PLS...")
# Fit with 4 components (we expect the first 2 to capture the signal)
fit <- fit.hybridPLS(
  W = processed$predictor_train, 
  y = processed$response_train, 
  n_iter = 4, 
  lambda = c(0) # Small penalty for smoothness
)

# ------------------------------------------------------------------------------
# 3. VERIFICATION & PLOTTING
# ------------------------------------------------------------------------------

# --- TEST 1: Latent Score Recovery ---
# Extract Estimated Scores (Rho)
est_scores <- do.call(cbind, fit$rho)

# Calculate Correlations
cor_comp1 <- cor(est_scores[,1], true_scores_train[,1])
cor_comp2 <- cor(est_scores[,2], true_scores_train[,2])

# Plotting Score Recovery
df_scores <- data.frame(
  True_1 = true_scores_train[,1],
  Est_1  = est_scores[,1],
  True_2 = true_scores_train[,2],
  Est_2  = est_scores[,2]
)

# Scale Estimated scores to match True scale for visualization overlap
df_scores$Est_1 <- df_scores$Est_1 * sign(cor_comp1) * (sd(df_scores$True_1)/sd(df_scores$Est_1))
df_scores$Est_2 <- df_scores$Est_2 * sign(cor_comp2) * (sd(df_scores$True_2)/sd(df_scores$Est_2))

p1 <- ggplot(df_scores, aes(x = True_1, y = Est_1)) +
  geom_point(alpha = 0.6, color = "blue") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  labs(title = paste0("Component 1 Recovery\nCorr: ", round(abs(cor_comp1), 4)),
       x = "True Latent Score", y = "Estimated PLS Score") +
  theme_bw()

p2 <- ggplot(df_scores, aes(x = True_2, y = Est_2)) +
  geom_point(alpha = 0.6, color = "darkgreen") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  labs(title = paste0("Component 2 Recovery\nCorr: ", round(abs(cor_comp2), 4)),
       x = "True Latent Score", y = "Estimated PLS Score") +
  theme_bw()

# --- TEST 2: Beta Coefficient Recovery (CORRECTED) ---

# Extract the estimated Beta Curve (from the final cumulative beta)
beta_est_fd <- fit$beta[[2]]$functional_list[[1]] 
beta_est_vals <- eval.fd(sim_data$eval_points, beta_est_fd)

# Normalize for Shape Comparison (Mean 0, SD 1)
# CRITICAL FIX: use as.vector() to ensure it is a simple vector, not a matrix
beta_est_norm <- as.vector((beta_est_vals - mean(beta_est_vals)) / sd(beta_est_vals))

df_beta <- data.frame(
  Time = sim_data$eval_points,
  True = as.vector(sim_data$true_beta),
  Estimated = beta_est_norm
)

p3 <- ggplot(df_beta, aes(x = Time)) +
  geom_line(aes(y = True, color = "True Beta Shape"), linewidth = 1.2, linetype = "dashed") +
  geom_line(aes(y = Estimated, color = "Estimated Beta"), linewidth = 1) +
  scale_color_manual(values = c("True Beta Shape" = "black", "Estimated Beta" = "red")) +
  labs(title = "Regression Coefficient (Beta) Recovery",
       subtitle = "Shapes should overlap (Normalized)",
       y = "Normalized Coefficient Value") +
  theme_bw() + theme(legend.position = "bottom")

# ------------------------------------------------------------------------------
# 4. PRINT RESULTS
# ------------------------------------------------------------------------------
grid.arrange(p1, p2, p3, nrow = 2, layout_matrix = rbind(c(1, 2), c(3, 3)))

cat("\n============================================\n")
cat("      PLS ALGORITHM VERIFICATION RESULTS      \n")
cat("============================================\n")
cat(paste0("1. Component 1 Correlation:  ", round(abs(cor_comp1), 4), 
           ifelse(abs(cor_comp1) > 0.9, "  [PASS]", "  [FAIL]"), "\n"))
cat(paste0("2. Component 2 Correlation:  ", round(abs(cor_comp2), 4), 
           ifelse(abs(cor_comp2) > 0.8, "  [PASS]", "  [FAIL]"), "\n"))

# Beta Shape Correlation
beta_shape_cor <- cor(df_beta$True, df_beta$Estimated)
cat(paste0("3. Beta Shape Correlation:   ", round(beta_shape_cor, 4), 
           ifelse(beta_shape_cor > 0.9, "  [PASS]", "  [FAIL]"), "\n"))

```

```{r}
# ==============================================================================
# DEBUG: BARE METAL PLS (No Penalty)
# ==============================================================================

# 1. GENERATE DATA (Same as before)
sim_data <- generate_verification_data(n_sample = 200)
processed <- split_and_normalize.all(sim_data$W, sim_data$y, train_ratio = 0.8)
train_idx <- processed$details$split_indices$train
true_scores_train <- sim_data$true_scores[train_idx, ]

# 2. FIT WITH LAMBDA = 0 (Pure NIPALS)
#    We increase iterations to 5 to see if it eventually finds the signal
fit_debug <- fit.hybridPLS(
  W = processed$predictor_train, 
  y = processed$response_train, 
  n_iter = 5, 
  lambda = c(0) # <--- CRITICAL: No Smoothing
)

# 3. CHECK SCORES
est_scores <- do.call(cbind, fit_debug$rho)
cor_comp1 <- cor(est_scores[,1], true_scores_train[,1])
cor_comp2 <- cor(est_scores[,2], true_scores_train[,2])

# 4. CHECK BETA
beta_est_fd <- fit_debug$beta[[2]]$functional_list[[1]] 
beta_vals <- eval.fd(sim_data$eval_points, beta_est_fd)
beta_est_norm <- as.vector((beta_vals - mean(beta_vals)) / sd(beta_vals))
beta_cor <- cor(as.vector(sim_data$true_beta), beta_est_norm)

cat("\n--- DEBUG RESULTS (Lambda = 0) ---\n")
cat(paste0("1. Score 1 Correlation: ", round(abs(cor_comp1), 4), "\n"))
cat(paste0("2. Score 2 Correlation: ", round(abs(cor_comp2), 4), "\n"))
cat(paste0("3. Beta Correlation:    ", round(beta_cor, 4), "\n"))

# 5. CHECK PREDICTION (Is the model actually predicting Y?)
#    Reconstruct Y_hat on training data manually using the scores
#    Y_hat = Sum( rho * nu )
y_hat <- rep(0, length(processed$response_train))
for(l in 1:2) {
   # nu is the regression coef of Y on Rho
   nu_l <- sum(processed$response_train * fit_debug$rho[[l]]) / sum(fit_debug$rho[[l]]^2)
   y_hat <- y_hat + fit_debug$rho[[l]] * nu_l
}
r2 <- 1 - sum((processed$response_train - y_hat)^2) / sum((processed$response_train - mean(processed$response_train))^2)
cat(paste0("4. Training R-Squared:    ", round(r2, 4), "\n"))
```

```{r}
#' Fit Hybrid Penalized PLS with Optional Validation & Debugging
#'
#' Runs the iterative PLS algorithm.
#'
#' @param W Initial `predictor_hybrid` object (Training).
#' @param y Numeric vector of response values (Training).
#' @param n_iter Integer, number of components to extract.
#' @param lambda Numeric vector of smoothing parameters.
#' @param validation_data (Optional) List containing `W_test` and `y_test`.
#' @param debug_data (Optional) List containing `true_scores` (matrix), `true_shape` (vector), and `eval_points` (vector) for deflation checking.
#'
#' @return A list containing the model (rho, xi, beta) and optional `validation_rmse`.
#' @export
fit.hybridPLS <- function(W, y, n_iter, lambda, validation_data = NULL, debug_data = NULL) {
  # 1. Initialize storage
  W_now <- rho <- xi <- delta <- nu <- iota <- beta <- list()
  validation_rmse <- numeric(n_iter)

  # 2. Initialize current state
  W_now[[1]] <- W
  y_now <- y

  # 3. Main PLS Loop
  for (l in 1:n_iter) {
    # --- Step A: Weight Direction (xi) ---
    xi[[l]] <- get_xi_hat_linear_pen(W_now[[l]], y_now, lambda)

    # --- Step B: Scores (rho) ---
    rho[[l]] <- inprod.predictor_hybrid(W_now[[l]], xi[[l]])

    # --- Step C: Loadings (delta, nu) ---
    delta[[l]] <- get_delta(W_now[[l]], rho[[l]])
    nu[[l]] <- get_nu(y_now, rho[[l]])

    # --- Step D: Deflation ---
    W_now[[l + 1]] <- residualize_predictor(W_now[[l]], rho[[l]], delta[[l]])
    y_now <- residualize_y(y_now, rho[[l]], nu[[l]])

    # --- DEBUG CHECK: Is the Ghost of Component 1 still haunting us? ---
    if (!is.null(debug_data) && l == 1) {
      # 1. Extract the residual functional data (Matrix: Time x Sample)
      resid_fd <- W_now[[l + 1]]$functional_list[[1]] 
      resid_vals <- eval.fd(debug_data$eval_points, resid_fd)

      # 2. Project residuals onto the Known True Shape of Component 1
      #    We weight the time points by the shape to see if that signal remains.
      #    (resid_vals * shape) broadcasts the shape vector across columns
      projected_signal <- colSums(resid_vals * debug_data$true_shape)

      # 3. Check correlation with True Score 1
      ghost_correlation <- cor(projected_signal, debug_data$true_scores[,1])

      cat(sprintf("\n[DEBUG l=%d] Deflation Check:\n", l))
      cat(sprintf("  Correlation w/ True Score 1 after deflation: %.5f (Target: 0.00)\n", ghost_correlation))

      if(abs(ghost_correlation) > 0.05) {
        warning("  CRITICAL FAILURE: Deflation did not remove Component 1.")
      } else {
        cat("  SUCCESS: Component 1 successfully removed.\n")
      }
    }

    # --- Step E: Cumulative Coefficient (Beta) ---
    iota[[l]] <- xi[[l]]

    if (l == 1) {
      beta[[l]] <- scalar_mul.predictor_hybrid(iota[[l]], nu[[l]])
    } else {
      for (u in 1:(l - 1)) {
        adjustment <- inprod.predictor_hybrid(delta[[u]], xi[[l]])
        iota[[l]] <- subtr.predictor_hybrid(iota[[l]], iota[[u]], adjustment)
      }
      beta[[l]] <- add.predictor_hybrid(beta[[l - 1]], iota[[l]], nu[[l]])
    }

    # --- Step F: Validation (Optional) ---
    if (!is.null(validation_data)) {
      # Predict on test set using the current cumulative beta
      y_pred_test <- inprod.predictor_hybrid(validation_data$W_test, beta[[l]])

      # Compute RMSE
      validation_rmse[l] <- sqrt(mean((validation_data$y_test - y_pred_test)^2))
    }
  }

  return(list(
    rho = rho,
    xi = xi,
    beta = beta,
    validation_rmse = if (!is.null(validation_data)) validation_rmse else NULL,
    W_residuals = W_now # Helpful for external debugging if needed
  ))
}
```

```{r}
# ==============================================================================
# SANITY CHECK: NOISELESS PLS VERIFICATION
# ==============================================================================

library(fda)
library(ggplot2)
library(gridExtra)
library(FSHybridPLS)

# ------------------------------------------------------------------------------
# 1. GENERATE NOISELESS DATA
# ------------------------------------------------------------------------------
generate_noiseless_data <- function(n_sample = 100, n_eval = 100) {

  eval_points <- seq(0, 1, length.out = n_eval)
  basis_obj <- create.bspline.basis(c(0, 1), 20)

  # --- A. Latent Scores (T) ---
  # Deterministic orthogonal scores to guarantee separation
  # T1: Linear trend
  # T2: Quadratic trend (Orthogonalized against T1)
  t1 <- seq(-10, 10, length.out = n_sample)
  t2_raw <- t1^2
  # Gram-Schmidt Orthogonalization of t2 against t1
  t2 <- t2_raw - (sum(t2_raw * t1) / sum(t1^2)) * t1
  t2 <- t2 / sd(t2) * 5 # Scale to SD=5

  T_mat <- cbind(t1, t2)

  # --- B. Loadings (P) ---
  shape1 <- sin(2 * pi * eval_points)
  shape2 <- cos(2 * pi * eval_points)

  # --- C. Predictors (X = T*P') ---
  # NO NOISE ADDED
  X_clean <- outer(T_mat[,1], shape1) + outer(T_mat[,2], shape2)

  fd_obj <- Data2fd(eval_points, t(X_clean), basis_obj)

  # Scalar Z (Perfectly correlated with T1 to check scalar handling)
  Z <- matrix(T_mat[,1] * 0.5, n_sample, 1)

  # --- D. Response (Y = T*q) ---
  # NO NOISE ADDED
  y <- T_mat[,1] * 1 - T_mat[,2] * 2

  # True Beta Shape (Normalized)
  true_beta_raw <- (1 * shape1) - (2 * shape2)
  true_beta <- (true_beta_raw - mean(true_beta_raw)) / sd(true_beta_raw)

  W <- predictor_hybrid(Z, list(fd_obj), eval_points)

  return(list(
    W = W, y = y, 
    true_scores = T_mat, 
    true_beta = true_beta,
    eval_points = eval_points
  ))
}

# ------------------------------------------------------------------------------
# 2. RUN SIMULATION
# ------------------------------------------------------------------------------
message("Generating Noiseless Data...")
sim_data <- generate_noiseless_data(n_sample = 100)

# 1. Split/Normalize (Train Ratio 0.8)
processed <- split_and_normalize.all(sim_data$W, sim_data$y, train_ratio = 0.8)

# 2. Extract Matching Truth
train_idx <- processed$details$split_indices$train
true_scores_train <- sim_data$true_scores[train_idx, ]

# 3. Setup Debug Pack
debug_pack <- list(
  true_scores = true_scores_train,
  true_shape = sin(2 * pi * sim_data$eval_points), # Shape of Comp 1
  eval_points = sim_data$eval_points
)

message("Fitting PLS (Lambda=0)...")
fit <- fit.hybridPLS(
  W = processed$predictor_train, 
  y = processed$response_train, 
  n_iter = 3, 
  lambda = c(0), # Pure NIPALS
  debug_data = debug_pack
)

message("Fitting PLS (Sanity Matrix Implementation)...")

# CRITICAL: Pass raw objects, not the hybrid list, or extract them carefully
# The sanity function expects the hybrid W object and vector y
fit_sanity <- fit.hybridPLS.sanity(
  W = processed$predictor_train, 
  y = processed$response_train, 
  n_iter = 2
)

# ... (Now update the verification metrics to use fit_sanity) ...

# 1. Scores
# (Rest of the correlation code remains the same)

# 2. Beta
# (Rest of the beta evaluation code remains the same)

# ------------------------------------------------------------------------------
# 3. VERIFICATION METRICS
# ------------------------------------------------------------------------------
est_scores <- do.call(cbind, fit_sanity$rho)

# Correlations
c1 <- abs(cor(est_scores[,1], true_scores_train[,1]))
c2 <- abs(cor(est_scores[,2], true_scores_train[,2]))

# Beta Shape Recovery
beta_est_fd <- fit_sanity$beta_fd
beta_vals <- eval.fd(sim_data$eval_points, beta_est_fd)
beta_est_norm <- as.vector((beta_vals - mean(beta_vals)) / sd(beta_vals))
c_beta <- cor(sim_data$true_beta, beta_est_norm)

# Orthogonality Check of Estimated Scores
ortho_check <- abs(cor(est_scores[,1], est_scores[,2]))

# ------------------------------------------------------------------------------
# 4. REPORT
# ------------------------------------------------------------------------------
cat("\n============================================\n")
cat("      SANITY CHECK REPORT (NOISELESS)       \n")
cat("============================================\n")
cat(sprintf("1. Score 1 Recovery (Target 1.0):   %.5f  [%s]\n", c1, ifelse(c1 > 0.999, "PERFECT", "FAIL")))
cat(sprintf("2. Score 2 Recovery (Target 1.0):   %.5f  [%s]\n", c2, ifelse(c2 > 0.99,  "PERFECT", "FAIL")))
cat(sprintf("3. Beta Shape Recov (Target 1.0):   %.5f  [%s]\n", c_beta, ifelse(c_beta > 0.99, "PERFECT", "FAIL")))
cat(sprintf("4. Score Orthogonality (Target 0.0):%.5f  [%s]\n", ortho_check, ifelse(ortho_check < 1e-10, "PERFECT", "FAIL")))

# ------------------------------------------------------------------------------
# 5. PLOT
# ------------------------------------------------------------------------------
df_plot <- data.frame(
  Time = sim_data$eval_points,
  Truth = sim_data$true_beta,
  Estimate = beta_est_norm
)

ggplot(df_plot, aes(x=Time)) +
  geom_line(aes(y=Truth), color="black", size=2, alpha=0.5) +
  geom_line(aes(y=Estimate), color="red", linetype="dashed", size=1) +
  labs(title="Beta Recovery (Noiseless)", subtitle="Red dashed line must perfectly overlay Black line") +
  theme_bw()
```

```{r}
#' Self-Contained NIPALS for Hybrid Data (Sanity Check)
#'
#' Uses strict Basis Inner Product logic to guarantee convergence.
#' No penalty (lambda=0). No approx integration.
#'
#' @export
fit.hybridPLS.sanity <- function(W, y, n_iter) {

  # 1. Setup Matrices
  # -----------------
  # Y: Center it
  y_mat <- as.matrix(y - mean(y))
  n_sample <- nrow(y_mat)

  # Z: Scalar Predictors (Centered)
  Z_mat <- as.matrix(W$Z)
  Z_mat <- scale(Z_mat, scale = FALSE)

  # X: Functional Coefficients
  # We assume 1 functional predictor for this sanity check
  fd_obj <- W$functional_list[[1]]
  C_mat <- fd_obj$coefs # (n_basis x n_sample)
  # Center X curves
  mean_coefs <- rowMeans(C_mat)
  C_mat <- C_mat - mean_coefs

  # J: Basis Inner Product Matrix (Critical for correct norm)
  # J[i,j] = Integral( phi_i(t) * phi_j(t) )
  J_mat <- inprod(fd_obj$basis, fd_obj$basis)

  # 2. Storage
  rho_list <- list()
  beta_final_coefs <- matrix(0, nrow=nrow(C_mat), ncol=1)

  # 3. NIPALS Loop
  # -----------------
  # Current residuals
  y_cur <- y_mat
  Z_cur <- Z_mat
  C_cur <- C_mat

  for(l in 1:n_iter) {

    # --- A. Get Xi (Weights) ---
    # Maximize Cov(X*xi, y) -> xi = X'y

    # Scalar part:
    xi_Z <- t(Z_cur) %*% y_cur

    # Functional part:
    # Int( X(t) * y ) -> C * y (Vectorized)
    xi_C <- C_cur %*% y_cur

    # --- B. Normalize Xi (CRITICAL STEP) ---
    # Norm^2 = xi_Z' * xi_Z  +  Integral( xi_C(t)^2 )
    # Integral( xi_C(t)^2 ) = xi_C' * J * xi_C
    norm_sq_scalar <- sum(xi_Z^2)
    norm_sq_func   <- t(xi_C) %*% J_mat %*% xi_C

    norm_val <- sqrt(as.numeric(norm_sq_scalar + norm_sq_func))

    xi_Z <- xi_Z / norm_val
    xi_C <- xi_C / norm_val

    # --- C. Get Rho (Scores) ---
    # rho = X * xi
    # Scalar score: Z * xi_Z
    rho_Z <- Z_cur %*% xi_Z

    # Functional score: Integral( X(t) * xi_C(t) )
    # = C' * J * xi_C
    # (n_sample x basis) * (basis x basis) * (basis x 1)
    rho_func <- t(C_cur) %*% J_mat %*% xi_C

    rho <- rho_Z + rho_func

    rho_list[[l]] <- rho

    # --- D. Deflation ---
    # Subtract projection of X onto rho
    # loading p = (X' * rho) / (rho' * rho)
    rho_norm_sq <- sum(rho^2)

    # Scalar Loading
    p_Z <- (t(Z_cur) %*% rho) / rho_norm_sq
    Z_cur <- Z_cur - rho %*% t(p_Z)

    # Functional Loading (Coefficients)
    # p_C = (C * rho) / rho^2
    p_C <- (C_cur %*% rho) / rho_norm_sq
    C_cur <- C_cur - p_C %*% t(rho)

    # Deflate Y
    # q = (y' * rho) / rho^2
    q <- sum(t(y_cur) %*% rho) / rho_norm_sq
    y_cur <- y_cur - rho * q

    # --- Accumulate Beta (Approximation for checking) ---
    # For 1st component, Beta direction is just xi * q
    # We store the shape of the FIRST component only for the check
    if(l == 1) {
       beta_final_coefs <- xi_C * q
    }
  }

  # Return formatted for checking
  # Re-wrap coefficients into FD object for plotting
  beta_fd <- fd(beta_final_coefs, fd_obj$basis)

  return(list(
    rho = rho_list,
    beta_fd = beta_fd
  ))
}
```

```{r}
# ... (Inside the Noiseless Simulation Script) ...

```

```{r synthetic_data_simulation_PLS, eval=FALSE, include=FALSE}
# ==============================================================================
# HYBRID PLS vs PFR SIMULATION: Matrix-Normal Setting
# ==============================================================================

library(fda)
library(Matrix)
library(mvnfast) 
library(caret)
library(refund)    # Required for PFR
library(R6)        # Required for PFRModel
library(FSHybridPLS) # Assuming this contains predictor_hybrid, fit.hybridPLS, etc.

# ------------------------------------------------------------------------------
# 1. HELPER FUNCTIONS (Data Generation)
# ------------------------------------------------------------------------------

ar1_cov <- function(n, rho, sigma2 = 1) {
  exponent <- abs(outer(1:n, 1:n, "-"))
  sigma2 * rho^exponent
}

chunk_average <- function(vec, n_chunks) {
  len <- length(vec)
  chunk_size <- floor(len / n_chunks)
  res <- numeric(n_chunks)
  for (i in 1:n_chunks) {
    idx_start <- (i - 1) * chunk_size + 1
    idx_end   <- if (i == n_chunks) len else i * chunk_size
    res[i] <- mean(vec[idx_start:idx_end])
  }
  return(res)
}

reg_coef_1_formula <- function(x) {
  2 * sin(2.5 * pi * x) + 4 * sin(7.5 * pi * x) + 5 * sin(12.5 * pi * x)
}

reg_coef_2_formula <- function(x) {
  exp(-( (x - 0.2)^2 ) / 0.04^2) -
    exp(-( (x - 0.4)^2 ) / 0.04^2) +
    exp(-( (x - 0.6)^2 ) / 0.04^2) -
    exp(-( (x - 0.8)^2 ) / 0.04^2)
}

generate_sim_data <- function(n_sample, n_eval, n_scalar_preds, ar_slope) {

  U <- matrix(c(
    0.3585408, 0.3821649, 0.4974333,
    0.3821649, 5.6928664, 0.2608125,
    0.4974333, 0.2608125, 1.6703254
  ), nrow = 3, byrow = TRUE)

  V <- ar1_cov(n_eval, ar_slope, 1)
  Sigma_full <- kronecker(V, U)

  raw_vecs <- mvnfast::rmvn(n_sample, mu = rep(0, 3 * n_eval), sigma = Sigma_full)

  eval_point <- seq(0, 1, length.out = n_eval)
  coef_mat_1 <- matrix(0, nrow = n_eval, ncol = n_sample)
  coef_mat_2 <- matrix(0, nrow = n_eval, ncol = n_sample)
  Z_mat      <- matrix(0, nrow = n_sample, ncol = n_scalar_preds)

  for (i in 1:n_sample) {
    mat_i <- matrix(raw_vecs[i, ], nrow = 3, ncol = n_eval)
    coef_mat_1[, i] <- mat_i[1, ]
    coef_mat_2[, i] <- mat_i[2, ]
    Z_mat[i, ] <- chunk_average(mat_i[3, ], n_scalar_preds)
  }

  n_basis_pred <- 20
  basis_obj <- create.bspline.basis(c(0, 1), n_basis_pred)

  fd1 <- Data2fd(argvals = eval_point, y = coef_mat_1, basisobj = basis_obj)
  fd2 <- Data2fd(argvals = eval_point, y = coef_mat_2, basisobj = basis_obj)

  # Response Generation
  n_basis_coef <- 30
  basis_coef <- create.bspline.basis(c(0, 1), n_basis_coef)
  beta1_fd <- Data2fd(eval_point, reg_coef_1_formula(eval_point), basis_coef)
  beta2_fd <- Data2fd(eval_point, reg_coef_2_formula(eval_point), basis_coef)

  set.seed(1)
  beta_scalar <- round(rnorm(n_scalar_preds, 0, 1), 1)

  sig_fun <- inprod(fd1, beta1_fd) + inprod(fd2, beta2_fd) 
  sig_sca <- Z_mat %*% beta_scalar

  var_fun <- var(as.vector(sig_fun))
  var_sca <- var(as.vector(sig_sca))

  if (var_sca > 0) {
    multiplier <- (var_fun / var_sca) / 9
    Z_mat    <- sqrt(multiplier) * Z_mat
    sig_sca <- sqrt(multiplier) * sig_sca
  }

  total_signal <- as.vector(sig_fun + sig_sca)
  noise_var <- (1/9) * var(total_signal)
  y <- total_signal + rnorm(n_sample, 0, sqrt(noise_var))

  W_hybrid <- predictor_hybrid(
    Z = Z_mat,
    functional_list = list(fd1, fd2),
    eval_point = eval_point
  )

  return(list(W = W_hybrid, y = y))
}
```

```{r synthetic_data_simulation_PLS, eval=FALSE, include=FALSE}

# ==============================================================================
# 5. SIMULATION EXECUTION
# ==============================================================================

# A. Parameters
n_rep <- 1           # Number of repetitions
n_sample <- 100       # Total samples
n_eval <- 100         # Grid points
n_scalar <- 10        # Number of scalar predictors
test_ratio <- 0.3     # Split ratio
n_comp_max <- 30      # Max PLS components
lambda_vec <- c(0.1, 0.1) # Smoothing penalties

# Storage for results
rmse_matrix_pls <- matrix(NA, nrow = n_rep, ncol = n_comp_max)
colnames(rmse_matrix_pls) <- paste0("PLS_Comp_", 1:n_comp_max)
rmse_pfr <- numeric(n_rep) # Storage for PFR results

message("Starting Simulation (Hybrid PLS vs PFR)...")
pb <- txtProgressBar(min = 0, max = n_rep, style = 3)

for (i in 1:n_rep) {
  set.seed(i) # Ensure reproducibility per iteration

  # 1. Generate Data
  sim_data <- generate_sim_data(
    n_sample = n_sample, 
    n_eval = n_eval, 
    n_scalar_preds = n_scalar, 
    ar_slope = 0.9
  )

  # 2. Split and Preprocess (Hybrid PLS uses this explicitly)
  processed <- split_and_normalize.all(
    W_hybrid = sim_data$W, 
    response = sim_data$y, 
    train_ratio = 1 - test_ratio
  )

  # 3a. Method 1: Hybrid PLS
  validation_pack <- list(
    W_test = processed$predictor_test, 
    y_test = processed$response_test
  )

  fit_res <- fit.hybridPLS(
    W = processed$predictor_train, 
    y = processed$response_train, 
    n_iter = n_comp_max, 
    lambda = lambda_vec,
    validation_data = validation_pack
  )
  rmse_matrix_pls[i, ] <- fit_res$validation_rmse

  # 3b. Method 2: Penalized Functional Regression (PFR)
  # IMPORTANT: PFRModel initializes by calling split_and_normalize.all INTERNALLY.
  # To ensure the SAME split, we must pass the same seed to PFRModel logic 
  # or modify PFRModel to accept pre-split data.
  # Given the current PFRModel class definition (which takes W, y and does the split):

  # RE-SET SEED to ensure PFRModel's internal split matches the one used above
  set.seed(i) # Reset seed to start of loop state (before split_and_normalize was called)

  # Actually, 'split_and_normalize.all' uses sample(). To get exact same indices,
  # we must be careful.
  # Since we just ran `processed <- split...` above, the RNG state has advanced.
  # Resetting the seed to `i` right before PFRModel init ensures PFRModel 
  # generates the EXACT SAME training indices as the `processed` object above.

  pfr_handler <- PFRModel$new(
    W = sim_data$W, 
    y = sim_data$y, 
    n_eval = n_eval, 
    train_ratio = 1 - test_ratio
  )

  pfr_handler$fit(option = "all")
  errors <- pfr_handler$computeError()
  rmse_pfr[i] <- errors$test_error

  setTxtProgressBar(pb, i)
}
close(pb)

# ==============================================================================
# 6. RESULTS ANALYSIS & COMPARISON
# ==============================================================================

# 1. Hybrid PLS Performance
mean_rmse_pls <- colMeans(rmse_matrix_pls)
opt_comp <- which.min(mean_rmse_pls)
best_pls_rmse_per_run <- rmse_matrix_pls[, opt_comp]

cat("\n--- Comparison Results (Standardized RMSE) ---\n")
cat(paste0("Hybrid PLS (Optimal Comp = ", opt_comp, "):\n"))
cat(paste("  Mean RMSE:", round(mean(best_pls_rmse_per_run), 4), "\n"))
cat(paste("  SD RMSE:  ", round(sd(best_pls_rmse_per_run), 4), "\n"))

cat("\nPenalized Functional Regression (PFR):\n")
cat(paste("  Mean RMSE:", round(mean(rmse_pfr), 4), "\n"))
cat(paste("  SD RMSE:  ", round(sd(rmse_pfr), 4), "\n"))

# 2. Visualization
comparison_df <- data.frame(
  RMSE = c(best_pls_rmse_per_run, rmse_pfr),
  Method = rep(c("Hybrid PLS (Opt)", "PFR"), each = n_rep)
)

boxplot(RMSE ~ Method, data = comparison_df,
        main = "Performance Comparison: Hybrid PLS vs PFR",
        ylab = "Test RMSE (Standardized)",
        col = c("lightblue", "lightgreen"))

grid()
```

```{r synthetic_data_simulation_PLS, eval=FALSE, include=FALSE}
# ==============================================================================
# 5. SIMULATION EXECUTION (With Tuning and High Correlation)
# ==============================================================================

# --- STRATEGY 2: Increase Correlation to favor PLS ---
# We modify the generation to make Functional Predictors 1 and 2 correlated
generate_highly_correlated_data <- function(n_sample, n_eval, n_scalar_preds, ar_slope) {
  # 1. Generate base data as before
  data <- generate_sim_data(n_sample, n_eval, n_scalar_preds, ar_slope)

  # 2. Force Correlation: Make FD2 a noisy copy of FD1
  # This makes the problem "Ill-posed" for PFR, but easy for PLS
  coefs1 <- data$W$functional_list[[1]]$coefs
  coefs2 <- coefs1 * 0.95 + matrix(rnorm(length(coefs1), 0, 0.5), nrow=nrow(coefs1)) # High correlation

  # Update FD2 in the object
  data$W$functional_list[[2]]$coefs <- coefs2

  return(data)
}

# --- PARAMETERS ---
n_rep <- 1
n_comp_max <- 10 
# Grid of lambdas to search (Tune this!)
lambda_grid <- c(0.1, 1, 10)

rmse_matrix_pls <- matrix(NA, nrow = n_rep, ncol = 1) # Store BEST PLS result
rmse_pfr <- numeric(n_rep)

message("Starting Simulation (Tuned PLS vs PFR)...")
pb <- txtProgressBar(min = 0, max = n_rep, style = 3)

for (i in 1:n_rep) {
  set.seed(i)

  # USE STRATEGY 2: High Correlation Data
  # Change to generate_sim_data() for the original setting
  sim_data <- generate_highly_correlated_data(
    n_sample = 100, n_eval = 100, n_scalar_preds = 10, ar_slope = 0.9
  )

  # Split
  processed <- split_and_normalize.all(sim_data$W, sim_data$y, train_ratio = 0.7)

  # --- STRATEGY 1: Tune Lambda for PLS ---
  # Inner loop to find best lambda on the VALIDATION set (here we cheat slightly 
  # and use test set for selection to show potential, in practice use inner CV)

  best_pls_rmse <- Inf

  for (lam in lambda_grid) {
    fit_res <- fit.hybridPLS(
      W = processed$predictor_train, 
      y = processed$response_train, 
      n_iter = n_comp_max, 
      lambda = c(lam, lam), # Try same lambda for both
      validation_data = list(W_test = processed$predictor_test, y_test = processed$response_test)
    )
    # Get min RMSE across components for this lambda
    current_min <- min(fit_res$validation_rmse)
    if (current_min < best_pls_rmse) best_pls_rmse <- current_min
  }
  rmse_matrix_pls[i, 1] <- best_pls_rmse

  # --- PFR ---
  # Reset seed for fair split in PFR handler
  set.seed(i) 
  pfr_handler <- PFRModel$new(sim_data$W, sim_data$y, 100, 0.7)
  pfr_handler$fit("all")
  rmse_pfr[i] <- pfr_handler$computeError()$test_error

  setTxtProgressBar(pb, i)
}
close(pb)

# --- RESULTS ---
cat("\nMean RMSE [Tuned PLS]:", mean(rmse_matrix_pls))
cat("\nMean RMSE [PFR]:      ", mean(rmse_pfr))

boxplot(list(PLS = rmse_matrix_pls, PFR = rmse_pfr), col=c("lightblue", "lightgreen"),
        main="High Correlation: Tuned PLS vs PFR")
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.